---
title: "Table - Beaver Biodiv Analysis"
author: Liz Anderson
format: html
---

Use data generated from elb_Scaling_BeaverBiodiv.qmd to make a table to include in thesis chapter 2

The table will have the full list of species that were detected, and the mean and species richness for both the beaver sites and control sites

Set up
```{r}
rm(list=ls())
library(here)
library(tidyverse)
```
Read in the data
```{r}
table_df <- read_csv(here("Data", "results_wild_mean_per_day.csv"))
```

```{r}
## I need the mean detections per day across the full treatment - one value for each species for each treatment.

table_df |>
  group_by(treatment, species) |>
  summarise(mean_det_day_treatment = mean(mean_detections_per_day),
            sd_det_day_treatment = sd(mean_detections_per_day)) |>
  view()

## Realized that when taking the mean here it is the mean of all the rows of that species, so for a species only detected at one site is is keeping that value, but taking the average by the number of sites


```

```{r}
# all_deployments <- c("AM0063", "AM0064", "AM0081", "AM0083", "AM0087", "AM0094", "AM0095", "AM0096", "AM0067", "AM0066", "AM0086", "AM0089", "AM0075", "AM0073", "AM0074", "AM0090")

# all_treatment = c("beaver", "beaver", "beaver", "beaver", "beaver", "beaver", "beaver", "beaver", "control", "control", "control", "control", "control", "control", "control", "control")



## use tidyr - this takes all possible pairings of deployment_num and species and gives them a row

filled_df <- table_df |>
  complete(deployment_num, species, fill = list(mean_detections_per_day = 0, sd_detections_per_day = 0))

filled_df <- filled_df |>
  mutate(treatment = if_else(deployment_num == "AM0067" | deployment_num == "AM0066" | deployment_num == "AM0086" | deployment_num == "AM0089" | deployment_num == "AM0075" | deployment_num == "AM0073" | deployment_num == "AM0074" | deployment_num == "AM0090" | deployment_num == "AM0065",
                             true = "control",
                             false = "beaver"))
```

Now take the mean detections per day per each species per treatment
```{r}
mean_det_treatment <- filled_df |>
  group_by(treatment, species) |>
  summarise(mean_det_day = mean(mean_detections_per_day),
            sd_det_day = sd(mean_detections_per_day))

```


Want to arrange by the species with highest mean detections in the beaver sites
```{r}
means_join <- mean_det_treatment |>
  select(-sd_det_day) |>
  pivot_wider(names_from = treatment, values_from = mean_det_day) |>
  arrange(desc(beaver)) |>
  rename(mean_beaver = "beaver",
         mean_control = "control")

sd_join <- mean_det_treatment |>
  select(-mean_det_day) |>
  pivot_wider(names_from = treatment, values_from = sd_det_day) |>
  rename(sd_beaver = "beaver",
         sd_control = "control")

joined <- left_join(means_join, sd_join, by = join_by(species))

joined <- joined |>
  relocate(sd_beaver, .after = mean_beaver)
```

Want the proportion of sites that the species was detected at as a column in table to
```{r}
## go back to filled_df

prop_sites <- filled_df |>
  mutate(present = if_else(mean_detections_per_day == 0.00000000,
                           true = 0,
                           false = 1)) |>
  group_by(treatment, species) |>
  summarise(prop_sites = mean(present)) |>
  pivot_wider(names_from = treatment, values_from = prop_sites) |>
  rename(prop_beaver = "beaver",
         prop_control = "control")
```

Now join the proportion data with the mean and sd dataframe
```{r}
joined_all <- left_join(joined, prop_sites, by = join_by(species))

joined_all <- joined_all |>
  relocate(prop_beaver, .after = sd_beaver)
```

Now the table is ready to write to csv

```{r}
write_csv(joined_all, here("Output", "mean_perspp_pertreatment_perday.csv"))
```


