---
title: "Initial code for beaver spillover biodiversity"
author: Liz Anderson
format: html
---



## Set up

```{r}
rm(list=ls())
library(here)
library(tidyverse)
library(vegan)
```


```{r}
## this is the data file after doing the following section of code! Don't need to run the next section of code for data set up anymore 

results_deploy_final <- read.csv(here("Data", "results_final.csv"))
```


### don't ned to run this code

Read in data (after running through the file `Merging_output.qmd`)

```{r}
results_deploy <- read.csv(here("Data/merged_output_deployment.csv"))
birdnet_only <- read.csv(here("Data/birdnet_results.csv"))
```

##### Dataset Modifictions

Make a variable to differentiate beaver sites and control sites
```{r}
table(results_deploy$deployment_num)

results_deploy <- results_deploy |>
  mutate(treatment = if_else(deployment_num == "AM0067" | deployment_num == "AM0066" | deployment_num == "AM0086" | deployment_num == "AM0089" | deployment_num == "AM0075" | deployment_num == "AM0073" | deployment_num == "AM0074" | deployment_num == "AM0090" | deployment_num == "AM0065",
                             true = "control",
                             false = "beaver")) |>
  relocate(treatment)

table(results_deploy$treatment)

results_deploy_final <- results_deploy |>
  filter(deployment_num != "AM0065") 
## note taking out this control site removed ~1,000 observations

table(results_deploy_final$treatment)
## note the extreme difference in observations between beaver sites and control sites
```

```{r}
## separate the scientific name and common name for making better plots
results_deploy_final <- results_deploy_final |>
  separate(species, 
           into = c("species_scientific", "species"),
           sep = "_")

## write this dataset as a file
# write.csv(results_deploy_final, here("Data", "results_final.csv"))

```
_end of code that doesn't need to run_


## Calculations

* Number of AM days
* Total minutes of recording
* Species detected and how many times total
* Species detected and how many times for each site
* Diversity indices for each site
* Diversity indices for each treatment group

### Number of AM days 

#### All sites
```{r}
## count the number of days
length(unique(results_deploy_final$date))

## number of detections (3 second clips that detected something) on each day
results_deploy_final |>
  count(date)

## total seconds of recording that detected something -- (sum the counts from num observations and multiply by 3)
results_deploy_final |>
  count(date) |>
  summarise(total_times = sum(n)) |>
  mutate(seconds_detecting = total_times * 3) |>
  mutate(minutes_detecting = seconds_detecting / 60) |>
  mutate(hours_detecting = minutes_detecting / 60)

```

#### beaver sites
```{r}
results_deploy_beaver <- results_deploy_final |>
  filter(treatment == "beaver")

## count the number of days
length(unique(results_deploy_beaver$date))

## number of detections (3 second clips that detected something) on each day
results_deploy_beaver |>
  count(date)

## total seconds of recording that detected something -- (sum the counts from num observations and multiply by 3)
results_deploy_beaver |>
  count(date) |>
  summarise(total_times = sum(n)) |>
  mutate(seconds_detecting = total_times * 3) |>
  mutate(minutes_detecting = seconds_detecting / 60) |>
  mutate(hours_detecting = minutes_detecting / 60)
```

#### control sites
```{r}
results_deploy_control <- results_deploy_final |>
  filter(treatment == "control")

## count the number of days
length(unique(results_deploy_control$date))

## number of detections (3 second clips that detected something) on each day
results_deploy_control |>
  count(date)

## total seconds of recording that detected something -- (sum the counts from num observations and multiply by 3)
results_deploy_control |>
  count(date) |>
  summarise(total_times = sum(n)) |>
  mutate(seconds_detecting = total_times * 3) |>
  mutate(minutes_detecting = seconds_detecting / 60) |>
  mutate(hours_detecting = minutes_detecting / 60)
```
Control sites had much fewer time detections and time detecting
Only a 5 day difference in the number of days that detections were made on

_Should calculate how long each detector was out and able to record for... not sure off top of my head how best to do this_


### Species detected

The species detected and how many times they were detected

#### All sites
```{r}
library(pander)
results_deploy_final |>
  group_by(species) |>
  summarise(num_instances_spp = n()) |>
  pander()

## the total number of species/sounds recorded across all sites
length(unique(results_deploy_final$species))
```

#### Beavers

```{r}
## species list and number of detections
results_deploy_beaver |>
  group_by(species) |>
  summarise(num_instances_spp = n()) |>
  pander()

length(unique(results_deploy_beaver$species))
```

```{r}
## species and number of detections at all beaver deployments (lumped)
spp_numrec_depl_beaver <- results_deploy_beaver |>
  group_by(deployment_num, species) |>
  summarise(num_rec = n())
spp_numrec_depl_beaver

## the species and counts of each species recorded at each deployment (separated)

## AM0063
results_deploy_beaver |>
  filter(deployment_num == "AM0063") |>
  group_by(species) |>
  summarise(num_rec = n())

## AM0064
results_deploy_beaver |>
  filter(deployment_num == "AM0064") |>
  group_by(species) |>
  summarise(num_rec = n())

## AM0081
results_deploy_beaver |>
  filter(deployment_num == "AM0081") |>
  group_by(species) |>
  summarise(num_rec = n())

## AM0083
results_deploy_beaver |>
  filter(deployment_num == "AM0083") |>
  group_by(species) |>
  summarise(num_rec = n())

## AM0087
results_deploy_beaver |>
  filter(deployment_num == "AM0087") |>
  group_by(species) |>
  summarise(num_rec = n())

## AM0094
results_deploy_beaver |>
  filter(deployment_num == "AM0094") |>
  group_by(species) |>
  summarise(num_rec = n())

## AM0095
results_deploy_beaver |>
  filter(deployment_num == "AM0095") |>
  group_by(species) |>
  summarise(num_rec = n())

## AM0096
results_deploy_beaver |>
  filter(deployment_num == "AM0096") |>
  group_by(species) |>
  summarise(num_rec = n())

```

#### Control

```{r}
## species list and number of detections
results_deploy_control |>
  group_by(species) |>
  summarise(num_instances_spp = n()) |>
  pander()

length(unique(results_deploy_control$species)) ## more than half the number of species detected at beaver sites
```


```{r}
## the species and counts of each species recorded at each control deployment (lumped)
spp_numrec_depl_control <- results_deploy_control |>
  group_by(deployment_num, species) |>
  summarise(num_rec = n())
spp_numrec_depl_control

## the species and counts of each species recorded at each control deployment (separated)

## AM0063
results_deploy_control |>
  filter(deployment_num == "AM0067") |>
  group_by(species) |>
  summarise(num_rec = n())

## AM0064
results_deploy_control |>
  filter(deployment_num == "AM0066") |>
  group_by(species) |>
  summarise(num_rec = n())

## AM0081
results_deploy_control |>
  filter(deployment_num == "AM0086") |>
  group_by(species) |>
  summarise(num_rec = n())

## AM0083
results_deploy_control |>
  filter(deployment_num == "AM0089") |>
  group_by(species) |>
  summarise(num_rec = n())

## AM0087
results_deploy_control |>
  filter(deployment_num == "AM0075") |>
  group_by(species) |>
  summarise(num_rec = n())

## AM0094
results_deploy_control |>
  filter(deployment_num == "AM0073") |>
  group_by(species) |>
  summarise(num_rec = n())

## AM0095
results_deploy_control |>
  filter(deployment_num == "AM0074") |>
  group_by(species) |>
  summarise(num_rec = n())

## AM0096
results_deploy_control |>
  filter(deployment_num == "AM0090") |>
  group_by(species) |>
  summarise(num_rec = n())

```
## Scaling

Notes:
* For this detector for species - total detections divided by total days detector out
* Could then multiply by 10 days to standardize if weird number ^
* Then can use this data for vegan -- controlling for effort

```{r}
results_deploy_final |>
  mutate(date = ymd(date)) |>
  group_by(deployment_num) |>
  summarise(first_rec = min(date),
            last_rec = max(date))

results_deploy_final <- results_deploy_final |>
  mutate(date = ymd(date)) |>
  group_by(deployment_num) |>
  mutate(first_rec = min(date),
         last_rec = max(date))

## mutate a new column with lubridate to get number of deployment days for each - then group and summarize for each detector and species 

results_deploy_final <- results_deploy_final |>
  group_by(deployment_num) |>
  mutate(recording_days = last_rec - first_rec)

glimpse(results_deploy_final)

## to look at these columns:
results_deploy_final |>
  relocate(first_rec, last_rec, recording_days)


## range of recording days:
results_deploy_final |>
  ggplot(aes(x = recording_days))+
  geom_histogram()

results_deploy_final |>
  group_by(treatment) |>
  summarise(
    min_days = min(recording_days),
    max_days = max(recording_days)
  ) |>
  pander()
## looks fairly equal between the treatments
```

```{r}
## group by detector and species and count the total number of detections
## then divide by the number of days the detector recorded for (recording_days)

## recording days needs to be numeric, not as a time

results_deploy_final <- results_deploy_final |>
  mutate(recording_days = as.numeric(recording_days)) |>
  relocate(recording_days)

detections_perdeploy_perspp <- results_deploy_final |>
  group_by(deployment_num, species, recording_days) |>
  summarise(num_detections = n())
  

## get the average number of detections per 10 days
scaled_detections <- detections_perdeploy_perspp |>
  mutate(scaled_num_detections = (num_detections / recording_days) * 10)
scaled_detections


### Join to main dataset for if need...
## reduce to just the columns needed
scaled_join_test <- scaled_detections |>
  select(deployment_num, species, scaled_num_detections)

## join that to the big dataset
left_join(results_deploy_final, scaled_join_test, by = c("deployment_num", "species"))
 # view()
```

### Table for poster
Species in each treatment and scaled detections
Ideally avg detections per day
```{r}
avg_detections_spp_table <- scaled_detections |>
  mutate(treatment = if_else(deployment_num == "AM0067" | deployment_num == "AM0066" | deployment_num == "AM0086" | deployment_num == "AM0089" | deployment_num == "AM0075" | deployment_num == "AM0073" | deployment_num == "AM0074" | deployment_num == "AM0090" | deployment_num == "AM0065",
                             true = "control",
                             false = "beaver")) |>
  relocate(treatment) |>
  select(c("treatment", "deployment_num", "species", "scaled_num_detections")) |>
  group_by(treatment, species) |>
  summarise(mean_detections_10days = mean(scaled_num_detections))

avg_detections_spp_table_wide <- avg_detections_spp_table |>
  pivot_wider(names_from = treatment, 
              values_from = mean_detections_10days)
 
avg_detections_spp_table_wide[is.na(avg_detections_spp_table_wide)] <- 0

avg_detections_spp_table_wide_rounded <- avg_detections_spp_table_wide |>
  mutate(across(c(beaver, control), round, digits = 3))

library(gt)

gt(data = avg_detections_spp_table_wide_rounded)




# avg_detections_spp_table_wide_rounded |>
#  pander()




```




#### Wide data
```{r}
## try pivoting wider with the scaled numbers
## also filtered out engine and dog here

results_wide_scaled <- scaled_detections |>
  select(-3, -4) |>
  filter(species != "Engine" & species != "Dog") |>
  pivot_wider(names_from = species,
              values_from = scaled_num_detections)

results_wide_scaled[is.na(results_wide_scaled)] <- 0

results_wide_scaled <- results_wide_scaled |>
  mutate(treatment = if_else(deployment_num == "AM0067" | deployment_num == "AM0066" | deployment_num == "AM0086" | deployment_num == "AM0089" | deployment_num == "AM0075" | deployment_num == "AM0073" | deployment_num == "AM0074" | deployment_num == "AM0090" | deployment_num == "AM0065",
                             true = "control",
                             false = "beaver")) |>
  relocate(treatment)

results_wide_scaled
```



```{r}
species_by_treatment_df <- results_deploy_final |>
  group_by(species, treatment) |>
  summarise(num_detections = n()) |> 
  pivot_wider(names_from = treatment,
              values_from = num_detections)
species_by_treatment_df

## would be good to remake with scaled data...

```


## Plots

### Beavers
```{r}
## work in progress

ggplot(data = spp_numrec_depl_beaver, aes(x = species, y = num_rec)) +
  geom_segment(aes(x = species, xend = species,
                   y = 0, yend = num_rec)) +
  geom_point() +
  coord_flip() +
  facet_wrap(~deployment_num, scales = "free") +
  theme_bw() +
  labs(x = "Number of recorded detections",
       y = "Species")
```

### Control Sites
```{r}
ggplot(data = spp_numrec_depl_control, aes(x = species, y = num_rec)) +
  geom_segment(aes(x = species, xend = species,
                   y = 0, yend = num_rec)) +
  geom_point() +
  coord_flip() +
  facet_wrap(~deployment_num, scales = "free") +
  theme_bw() +
  labs(x = "Number of recorded detections",
       y = "Species")
```


## Species Accumulation Curve




## Next steps:
Species accumulation curve
- pivoting wider...
- assigning day numbers?
- figure out specific table

Calculate biodiversity indices with vegan

Comparison metrics and models with biodiversity index as response

Add a column for beaver sites about active beavers yes/no

Add in other spatial data from GIS - distance to development, distance to water


## 0.8 probability 
__Go see qmd for this analysis specifically__

Filter to just the rows where the probability is greather than or equal to 0.8
```{r}
# highprob_results_deploy <- results_deploy_final |>
#   filter(probability >= 0.8)
```

Check the species that were present and at which treatment with grouped and pivoted dataframe:
```{r}
## 0.8 and up probability
# highprob_by_treatment_df <- highprob_results_deploy |>
#  group_by(species, treatment) |>
#  summarise(num_detections = n()) |>
#  pivot_wider(names_from = treatment,
#              values_from = num_detections)


## compared to 0.4 and up probability
# species_by_treatment_df
# highprob_by_treatment_df
```

```{r}
# highprob_plot_df <- highprob_results_deploy |>
#  group_by(deployment_num, species, treatment) |>
#  summarise(num_rec = n())
#highprob_plot_df

#ggplot(data = highprob_plot_df, aes(x = species, y = num_rec, colour = treatment)) +
#  geom_segment(aes(x = species, xend = species,
#                   y = 0, yend = num_rec)) +
#  geom_point() +
#  coord_flip() +
#  facet_wrap(~deployment_num, scales = "free") +
#  theme_bw() +
#  labs(x = "Number of recorded detections",
#       y = "Species")
```

## plot detector date ranges

```{r}
## make a column for the numeric day of year
results_deploy_final$date <- as.Date(results_deploy_final$date)
results_deploy_final

results_deploy_final$day_of_year <- yday(results_deploy_final$date)
results_deploy_final
```




```{r}
ggplot(data = results_deploy_final, aes(x = date, color = treatment)) +
  geom_bar() +
  coord_flip() +
  facet_grid(~deployment_num) +
  theme_minimal()
```
First day of a recording 9/21 (sept 21) 
to last day of a recording 12/12 (dec 12)

Recording range for each camera:
```{r}
### this code was already done up above

# results_deploy_final |>
#  group_by(deployment_num) |>
#  summarise(first_rec = min(date),
#            last_rec = max(date))

# results_deploy_final <- results_deploy_final |>
#  group_by(deployment_num) |>
#  mutate(first_rec = min(date),
#         last_rec = max(date))

## mutate a new column with lubridate to get number of deployment days for each - then group and summarize for each detector and species 
```



## Naive occupancy
proportion of stations where species detected relative to the total number of stations sampled
value ranges 0 to 1

good to talk about if the beaver sites are similar or different to each other

 do multiple - across all, across beaver, across control for each species to see differences



## Ordination work - following along with Ordination-in-class R project
```{r}
results_wide_scaled
# results_scaled_vegan

```

```{r}
### ordination

## replacing results_wide with the scaled version - results_wide_scaled

## The following code block creates an object by running the `metaMDS` function in the `vegan` package. This function conducts a NMDS ordination.

BirdOrd <- metaMDS(
  results_wide_scaled[, 3:38], #tells vegan which columns just have species info
  distance = "bray", #tells vegan what kind of distance to calculate
  k = 2, #tells vegan that we want 2-d output
  trymax = 50) #tells vegan the number of tries


## Stress is a measure of how well our two-dimensional representation captures the information in our 16-dimensional space (the 16 species). In general, stress values need to be less than 0.2 to be acceptable.

BirdOrd$stress ## check if the stress value is less than 0.2
stressplot(BirdOrd)

gof <- goodness(BirdOrd) #goodness of fit
plot(BirdOrd, type = "t", main = "goodness of fit")
points(BirdOrd, display = "sites", cex=gof*100) ## this didn't work

# PlotInfo
```

Plot results further
```{r}
#| error: true
## The first step is to use a function called `fortify` to rearrange the output into a long-format dataframe familiar to ggplot2. Then we can make a basic plot. Then we will spruce up our basic plot.

ForPlotting <- fortify(BirdOrd) ## ran into issue here too
```
```{r}
#| error: true
## Now lets repair the labels in ForPlotting to represent deployment names
ForPlotting$label[1:8] <- PlotInfo$deployment_num
```

Now we will make our plots from the new ForPlotting dataframe.

Basic plot:
```{r}
#| error: true
ggplot(ForPlotting, aes(nmds1,nmds2,group = score))+
   geom_text(aes(label=label, color = score))+
   scale_color_manual(values = c(sites = "brown", species = "green"))+
   geom_point(aes(color=score, shape =score))+
   scale_color_manual(values = c(sites = "red", species = "black"))
```
Improved plot:
```{r}
#| error: true
ggplot()+
   geom_point(data = subset(ForPlotting, score =="sites"),
              mapping = aes(x = NMDS1, y = NMDS2),
              alpha = 0.8)+
   geom_segment(data=subset(ForPlotting, score == "species"),
                mapping = aes(x = 0, y = 0, xend = NMDS1,   yend=NMDS2),
                 arrow = arrow(length = unit(0.015, "npc"),
               type = "closed"),
                 color = "darkgray",
                 size = 1,
               alpha = 0.8)+
   geom_text(data = subset(ForPlotting, score == "species"),
             mapping = aes(label = label, x = NMDS1*1.1, y=NMDS2*1.1), alpha = 0.8)+
   geom_abline(intercept = 0, slope = 0, linetype = "dashed", size = 0.8, color = "gray")+
   geom_vline(aes(xintercept = 0), linetype = "dashed", size = 0.8, color = "gray")+
   theme(
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      panel.background = element_blank(),
      axis.line = element_line(color = "black"),
      legend.position = "none")
```

This is code up to line 181 in the intro-ordination-example.qmd file -- haven't run yet because of issues with early code.





## Calculate diversity indices
Use the species_wide_scaled dataset that was created up above in the "Scaling" section
```{r}
library(vegan) #turns on package for calculating diversity indices
library(betapart) #turns on package for beta diversity (needed for Jaccard)
library(rstatix) #turns on package with some matrix maniuplation functions

## finish setting up my dataframe to use in vegan (remove the extra columns)
Div_Indices <- results_wide_scaled[,1:2]

results_scaled_vegan <- results_wide_scaled[,-(1:2)]
results_scaled_vegan

## Species Richness
Div_Indices$SR <- specnumber(results_scaled_vegan)
Div_Indices

## Simpson's Index
Div_Indices$Simpson <- diversity(results_scaled_vegan, index = "simpson")
Div_Indices

## Shannon Index
Div_Indices$Shannon <- diversity(results_scaled_vegan, index = "shannon")
Div_Indices

## summary statistics
# filter to beaver and control so can be compared
Div_Indices_Beaver <- Div_Indices |>
  filter(treatment == "beaver")

Div_Indices_Control <- Div_Indices |>
  filter(treatment == "control")

Div_Ind_Summary_all <- do.call(cbind, lapply(Div_Indices, summary))
Div_Ind_Summary_all

Div_Ind_Summary_beaver <- do.call(cbind, lapply(Div_Indices_Beaver, summary))
Div_Ind_Summary_beaver

Div_Ind_Summary_control <- do.call(cbind, lapply(Div_Indices_Control, summary))
Div_Ind_Summary_control

Div_Ind_Summary_beaver ## Median values are higher for all indices of the beaver sites
Div_Ind_Summary_control
```
Visualize the diversity indices:
```{r}
## pivot to longer format
Div_Long <- Div_Indices |>
  pivot_longer(cols = SR:Shannon, names_to = "Index", values_to = "Value")


## plot
full_div_plot <- ggplot(Div_Long, aes(x = fct_reorder(deployment_num, treatment), y = Value, group = Index, color = treatment))+
  geom_point(size = 2) +
  facet_wrap(~Index, scales = "free") +
  labs(x = "Deployment number", color = "Treatment") +
  coord_flip() +
  theme_bw()
full_div_plot

# ggsave(here("Output", "full_div_plot.jpg"), plot = full_div_plot, width = 10, height = 7, units = "in")
```

```{r}
Div_Indices |>
  group_by(treatment) |>
  summarise(mean_SR = mean(SR),
            mean_Shannon = mean(Shannon),
            mean_Simpson = mean(Simpson))

SR_plot <- ggplot(data = Div_Indices, aes(x = treatment, y = SR)) +
  geom_boxplot()+
  geom_point() +
  labs(x = "Treatment", y = "Species Richness") +
  theme_bw()

Simpson_plot <- ggplot(data = Div_Indices, aes(x = treatment, y = Simpson)) +
  geom_boxplot()+
  geom_point() +
  labs(x = "Treatment", y = "Simpson's Index") +
  theme_bw()

Shannon_plot <- ggplot(data = Div_Indices, aes(x = treatment, y = Shannon)) +
  geom_boxplot()+
  geom_point() +
  labs(x = "Treatment", y = "Shannon Index") +
  theme_bw()

SR_plot
Simpson_plot
Shannon_plot

# ggsave(here("Output", "SR_Plot.jpg"), plot = SR_plot, width = 6, height = 11, units = "in")

# ggsave(here("Output", "Simpson_Plot.jpg"), plot = Simpson_plot, width = 6, height = 11, units = "in")

# ggsave(here("Output", "Shannon_Plot.jpg"), plot = Shannon_plot, width = 6, height = 11, units = "in")
```



### Jaccard index

```{r}
Jaccard <- betadiver(results_scaled_vegan, "j") #j indicates to use the a/a+b+c formula
#result is of class 'distance' - convert to a matrix
Jaccard <- as.matrix(Jaccard)
```

_Determine the shared species between sites:_

Now use `betapart`package to get shared species between sites. The function requires a set of presence/absence rather than abundance data so we have to convert our DF dataset to that.

```{r}
#requires a dataset of presence/absence data so make that first
PresAb<-results_scaled_vegan
	
PresAb[][PresAb[] != 0]<- 1 #assigns all values > 0  to 1 
PresAb
```
Now we can run the function to get the number of shared species between each pair of plots.

```{r}
library(betapart)
Beta<-betapart.core(PresAb)#result is a list; we just need one part
Shared<-Beta$shared
## this is little a, the number of species shared between the communities
```

_Create our display matrix_

Now it would be ideal if we could create a matrix with number of shared species ("a") above the diagonal, species richness of each plot along the diagonal, and Jaccard index below the diagonal.

Let's start by adding row and column names to our existing matrices to keep track of things.

```{r}
row.names(Jaccard)<-Div_Indices$deployment_num
colnames(Jaccard)<-Div_Indices$deployment_num

row.names(Shared)<-Div_Indices$deployment_num
colnames(Shared)<-Div_Indices$deployment_num

## adds row and column names so can see which communities are being compared
```

Since we want the number of shared species above the diagonal, let's start with the shared matrix. Conveniently, it also has the number of species per plot as its diagonal.

So all we really need to do is put the Jaccard data below the diagonal.

```{r}
#using base r
Shared[lower.tri(Shared)] <- Jaccard[lower.tri(Jaccard, diag = FALSE)]

## puts jaccard below the diagonal

Shared
```

Write matrix and diversity indices to a csv:
```{r}
# write.csv(Shared, here("Output","AudioDivMatrix.csv"), row.names = F)
## row.names=F stops an extra column with row names


## div indices
# write.csv(Div_Indices, here("Output", "AudioDiversity.csv"), row.names = F)
```

## t-tests

```{r}
# "t.test(Ozone ~ Garden.location, data = ozone)
# t.test(response variable ~ predictor variable, data = dataframe)"

t.test(Shannon ~ treatment, data = Div_Indices) ## signif!

t.test(Simpson ~ treatment, data = Div_Indices) ## signif!

t.test(SR ~ treatment, data = Div_Indices) ## signif!
```

## Landcover data

```{r}
lulc <- read.csv(here("Data", "LULC_spillover_diversity.csv"))
glimpse(lulc)
```
To do:
* Convert to long
* filter to just the percentages

```{r}
lulc_long <- lulc |>
  select(1, 2, 17:29) |>
  pivot_longer(cols = 3:15, names_to = "land_type", values_to = "percent_area")
```
Join to Div_Indices dataframe so that can be used in a model with div indices as response
```{r}
Div_Indices_joined <- left_join(Div_Indices, lulc_long, by = join_by(deployment_num == DEPLOYMENT_NUM))
Div_Indices
```

_Stat tests to run:_

* test to see if diff in div indices is significantly different -- paired t-test...
* multiple regression model to see what the best predictors of div indices are
   - predictors: percent of land type, beaver vs control, active beavers or not, distance to ___
   

## paired t-test

```{r}
# "t.test(Ozone ~ Garden.location, data = ozone)
# t.test(response variable ~ predictor variable, data = dataframe)"

t.test(Shannon ~ treatment, data = Div_Indices) ## signif!

t.test(Simpson ~ treatment, data = Div_Indices) ## signif!

t.test(SR ~ treatment, data = Div_Indices) ## signif!
```









