---
title: "Initial code for beaver spillover biodiversity"
author: Liz Anderson
format: html
---

Goals for this file:

* Think about and start results to report in paper
* Think about and start plots to report in paper


## Set up

```{r}
rm(list=ls())
library(here)
library(tidyverse)
```


Read in data (after running through the file `Merging_output.qmd`)

```{r}
results_deploy <- read.csv(here("Data/merged_output_deployment.csv"))
birdnet_only <- read.csv(here("Data/birdnet_results.csv"))
```

## Dataset Modifictions

Make a variable to differentiate beaver sites and control sites
```{r}
table(results_deploy$deployment_num)

results_deploy <- results_deploy |>
  mutate(treatment = if_else(deployment_num == "AM0067" | deployment_num == "AM0066" | deployment_num == "AM0086" | deployment_num == "AM0089" | deployment_num == "AM0075" | deployment_num == "AM0073" | deployment_num == "AM0074" | deployment_num == "AM0090" | deployment_num == "AM0065",
                             true = "control",
                             false = "beaver")) |>
  relocate(treatment)

table(results_deploy$treatment)

results_deploy_final <- results_deploy |>
  filter(deployment_num != "AM0065") 
## note taking out this control site removed ~1,000 observations

table(results_deploy_final$treatment)
## note the extreme difference in observations between beaver sites and control sites
```

```{r}
## separate the scientific name and common name for making better plots
results_deploy_final <- results_deploy_final |>
  separate(species, 
           into = c("species_scientific", "species"),
           sep = "_")
```


## Calculations

* Number of AM days
* Total minutes of recording
* Species detected and how many times total
* Species detected and how many times for each site
* Diversity indices for each site
* Diversity indices for each treatment group

### Number of AM days 

#### All sites
```{r}
## count the number of days
length(unique(results_deploy_final$date))

## number of detections (3 second clips that detected something) on each day
results_deploy_final |>
  count(date)

## total seconds of recording that detected something -- (sum the counts from num observations and multiply by 3)
results_deploy_final |>
  count(date) |>
  summarise(total_times = sum(n)) |>
  mutate(seconds_detecting = total_times * 3) |>
  mutate(minutes_detecting = seconds_detecting / 60) |>
  mutate(hours_detecting = minutes_detecting / 60)

```

#### beaver sites
```{r}
results_deploy_beaver <- results_deploy_final |>
  filter(treatment == "beaver")

## count the number of days
length(unique(results_deploy_beaver$date))

## number of detections (3 second clips that detected something) on each day
results_deploy_beaver |>
  count(date)

## total seconds of recording that detected something -- (sum the counts from num observations and multiply by 3)
results_deploy_beaver |>
  count(date) |>
  summarise(total_times = sum(n)) |>
  mutate(seconds_detecting = total_times * 3) |>
  mutate(minutes_detecting = seconds_detecting / 60) |>
  mutate(hours_detecting = minutes_detecting / 60)
```

#### control sites
```{r}
results_deploy_control <- results_deploy_final |>
  filter(treatment == "control")

## count the number of days
length(unique(results_deploy_control$date))

## number of detections (3 second clips that detected something) on each day
results_deploy_control |>
  count(date)

## total seconds of recording that detected something -- (sum the counts from num observations and multiply by 3)
results_deploy_control |>
  count(date) |>
  summarise(total_times = sum(n)) |>
  mutate(seconds_detecting = total_times * 3) |>
  mutate(minutes_detecting = seconds_detecting / 60) |>
  mutate(hours_detecting = minutes_detecting / 60)
```
Control sites had much fewer time detections and time detecting
Only a 5 day difference in the number of days that detections were made on

_Should calculate how long each detector was out and able to record for... not sure off top of my head how best to do this_


### Species detected

The species detected and how many times they were detected

#### All sites
```{r}
library(pander)
results_deploy_final |>
  group_by(species) |>
  summarise(num_instances_spp = n()) |>
  pander()

## the total number of species/sounds recorded across all sites
length(unique(results_deploy_final$species))
```

#### Beavers

```{r}
## species list and number of detections
results_deploy_beaver |>
  group_by(species) |>
  summarise(num_instances_spp = n()) |>
  pander()

length(unique(results_deploy_beaver$species))
```

```{r}
## species and number of detections at all beaver deployments (lumped)
spp_numrec_depl_beaver <- results_deploy_beaver |>
  group_by(deployment_num, species) |>
  summarise(num_rec = n())
spp_numrec_depl_beaver

## the species and counts of each species recorded at each deployment (separated)

## AM0063
results_deploy_beaver |>
  filter(deployment_num == "AM0063") |>
  group_by(species) |>
  summarise(num_rec = n())

## AM0064
results_deploy_beaver |>
  filter(deployment_num == "AM0064") |>
  group_by(species) |>
  summarise(num_rec = n())

## AM0081
results_deploy_beaver |>
  filter(deployment_num == "AM0081") |>
  group_by(species) |>
  summarise(num_rec = n())

## AM0083
results_deploy_beaver |>
  filter(deployment_num == "AM0083") |>
  group_by(species) |>
  summarise(num_rec = n())

## AM0087
results_deploy_beaver |>
  filter(deployment_num == "AM0087") |>
  group_by(species) |>
  summarise(num_rec = n())

## AM0094
results_deploy_beaver |>
  filter(deployment_num == "AM0094") |>
  group_by(species) |>
  summarise(num_rec = n())

## AM0095
results_deploy_beaver |>
  filter(deployment_num == "AM0095") |>
  group_by(species) |>
  summarise(num_rec = n())

## AM0096
results_deploy_beaver |>
  filter(deployment_num == "AM0096") |>
  group_by(species) |>
  summarise(num_rec = n())

```

#### Control

```{r}
## species list and number of detections
results_deploy_control |>
  group_by(species) |>
  summarise(num_instances_spp = n()) |>
  pander()

length(unique(results_deploy_control$species)) ## more than half the number of species detected at beaver sites
```


```{r}
## the species and counts of each species recorded at each beaver deployment (lumped)
spp_numrec_depl_control <- results_deploy_control |>
  group_by(deployment_num, species) |>
  summarise(num_rec = n())
spp_numrec_depl_control

## the species and counts of each species recorded at each deployment (separated)

## AM0063
results_deploy_control |>
  filter(deployment_num == "AM0067") |>
  group_by(species) |>
  summarise(num_rec = n())

## AM0064
results_deploy_control |>
  filter(deployment_num == "AM0066") |>
  group_by(species) |>
  summarise(num_rec = n())

## AM0081
results_deploy_control |>
  filter(deployment_num == "AM0086") |>
  group_by(species) |>
  summarise(num_rec = n())

## AM0083
results_deploy_control |>
  filter(deployment_num == "AM0089") |>
  group_by(species) |>
  summarise(num_rec = n())

## AM0087
results_deploy_control |>
  filter(deployment_num == "AM0075") |>
  group_by(species) |>
  summarise(num_rec = n())

## AM0094
results_deploy_control |>
  filter(deployment_num == "AM0073") |>
  group_by(species) |>
  summarise(num_rec = n())

## AM0095
results_deploy_control |>
  filter(deployment_num == "AM0074") |>
  group_by(species) |>
  summarise(num_rec = n())

## AM0096
results_deploy_control |>
  filter(deployment_num == "AM0090") |>
  group_by(species) |>
  summarise(num_rec = n())

```

#### Comparison

```{r}
results_deploy_final |>
  group_by(species, treatment) |>
  summarise(num_detections = n())
## this was not quite what I wanted... 

# see half piece of paper for dataframe wanted (columns: spp, beaver sites, control sites)
```




## Plots

### Beavers
```{r}
## work in progress

ggplot(data = spp_numrec_depl_beaver, aes(x = species, y = num_rec)) +
  geom_segment(aes(x = species, xend = species,
                   y = 0, yend = num_rec)) +
  geom_point() +
  coord_flip() +
  facet_wrap(~deployment_num, scales = "free") +
  theme_bw() +
  labs(x = "Number of recorded detections",
       y = "Species")
```

### Control Sites
```{r}
ggplot(data = spp_numrec_depl_control, aes(x = species, y = num_rec)) +
  geom_segment(aes(x = species, xend = species,
                   y = 0, yend = num_rec)) +
  geom_point() +
  coord_flip() +
  facet_wrap(~deployment_num, scales = "free") +
  theme_bw() +
  labs(x = "Number of recorded detections",
       y = "Species")
```


## Species Accumulation Curve




## Next steps:
Species accumulation curve
- pivoting wider...
- assigning day numbers?
- figure out specific table

Calculate biodiversity indices with vegan

Comparison metrics and models with biodiversity index as response

Add a column for beaver sites about active beavers yes/no

Add in other spatial data from GIS - distance to development, distance to water







