---
title: "Organized - Beaver Biodiv Analysis - Descriptive"
author: Liz Anderson
format: html
---

# Descriptive stats with the full data set
* How long each AM recorded for at each treatment
* What species were detected at each treatment and how many times (counts)


```{r}
rm(list=ls())
library(here)
library(tidyverse)
library(vegan)

theme_teach<-function(){
  theme(
  #modify text, axis
  axis.text = element_text(color = "black", size = 12),
  axis.title = element_text(color = "black", size = 16)
  )
}
```

```{r}
results_deploy_final <- read_csv(here("Data", "results_final.csv"))
```

## Basic descriptive stats

### Number of AM days 

#### All sites
```{r}
## count the number of days
length(unique(results_deploy_final$date))

## number of detections (3 second clips that detected something) on each day
results_deploy_final |>
  count(date)

## total seconds of recording that detected something -- (sum the counts from num observations and multiply by 3)
results_deploy_final |>
  count(date) |>
  summarise(total_times = sum(n)) |>
  mutate(seconds_detecting = total_times * 3) |>
  mutate(minutes_detecting = seconds_detecting / 60) |>
  mutate(hours_detecting = minutes_detecting / 60)

```


#### beaver sites
```{r}
results_deploy_beaver <- results_deploy_final |>
  filter(treatment == "beaver")

## count the number of days
length(unique(results_deploy_beaver$date))

## number of detections (3 second clips that detected something) on each day
results_deploy_beaver |>
  count(date)

## total seconds of recording that detected something -- (sum the counts from num observations and multiply by 3)
results_deploy_beaver |>
  count(date) |>
  summarise(total_times = sum(n)) |>
  mutate(seconds_detecting = total_times * 3) |>
  mutate(minutes_detecting = seconds_detecting / 60) |>
  mutate(hours_detecting = minutes_detecting / 60)
```

#### control sites
```{r}
results_deploy_control <- results_deploy_final |>
  filter(treatment == "control")

## count the number of days
length(unique(results_deploy_control$date))

## number of detections (3 second clips that detected something) on each day
results_deploy_control |>
  count(date)

## total seconds of recording that detected something -- (sum the counts from num observations and multiply by 3)
results_deploy_control |>
  count(date) |>
  summarise(total_times = sum(n)) |>
  mutate(seconds_detecting = total_times * 3) |>
  mutate(minutes_detecting = seconds_detecting / 60) |>
  mutate(hours_detecting = minutes_detecting / 60)
```

Control sites had much fewer time detections and time detecting
Only a 5 day difference in the number of days that detections were made on


### plot detector date ranges

```{r}
glimpse(results_deploy_final)

results_deploy_final$day_of_year <- yday(results_deploy_final$date)
## make a column for day of the year (julien date)
results_deploy_final
```

```{r}
## want to facet this in a good order
results_deploy_final <- results_deploy_final |>
  mutate(deployment_num = fct_reorder(deployment_num, treatment)) ## didn't work

ggplot(data = results_deploy_final, aes(x = date, color = treatment)) +
  geom_bar() +
  coord_flip() +
  facet_grid(~deployment_num) +
  theme_minimal() +
  labs(x = "Date", y = "Count of detections", color = "Treatment")
```

First day of a recording 9/21 (sept 21) 
to last day of a recording 12/12 (dec 12)

Recording range for each camera:
```{r}
results_deploy_final |>
  group_by(deployment_num) |>
  summarise(first_rec = min(date),
            last_rec = max(date)) 

results_deploy_final |>
  group_by(treatment) |>
  summarise(first_rec = min(date),
            last_rec = max(date)) |>
  group_by(treatment) |>
  summarise(total_days = last_rec - first_rec)



 results_deploy_final <- results_deploy_final |>
  group_by(deployment_num) |>
  mutate(first_rec = min(date),
         last_rec = max(date))

## mutate a new column with lubridate to get number of deployment days for each - then group and summarize for each detector and species 
```




### Species detected

The species detected and how many times they were detected

#### All sites
```{r}
library(pander)
results_deploy_final |>
  group_by(species) |>
  summarise(num_instances_spp = n()) |>
  pander()

## the total number of species/sounds recorded across all sites
length(unique(results_deploy_final$species))
```

#### Beavers

```{r}
## species list and number of detections
results_deploy_beaver |>
  group_by(species) |>
  summarise(num_instances_spp = n()) |>
  pander()

length(unique(results_deploy_beaver$species))
```

```{r}
## species and number of detections at all beaver deployments (lumped)
spp_numrec_depl_beaver <- results_deploy_beaver |>
  group_by(deployment_num, species) |>
  summarise(num_rec = n())
spp_numrec_depl_beaver

## the species and counts of each species recorded at each deployment (separated)

## AM0063
results_deploy_beaver |>
  filter(deployment_num == "AM0063") |>
  group_by(species) |>
  summarise(num_rec = n())

## AM0064
results_deploy_beaver |>
  filter(deployment_num == "AM0064") |>
  group_by(species) |>
  summarise(num_rec = n())

## AM0081
results_deploy_beaver |>
  filter(deployment_num == "AM0081") |>
  group_by(species) |>
  summarise(num_rec = n())

## AM0083
results_deploy_beaver |>
  filter(deployment_num == "AM0083") |>
  group_by(species) |>
  summarise(num_rec = n())

## AM0087
results_deploy_beaver |>
  filter(deployment_num == "AM0087") |>
  group_by(species) |>
  summarise(num_rec = n())

## AM0094
results_deploy_beaver |>
  filter(deployment_num == "AM0094") |>
  group_by(species) |>
  summarise(num_rec = n())

## AM0095
results_deploy_beaver |>
  filter(deployment_num == "AM0095") |>
  group_by(species) |>
  summarise(num_rec = n())

## AM0096
results_deploy_beaver |>
  filter(deployment_num == "AM0096") |>
  group_by(species) |>
  summarise(num_rec = n())

```

#### Control

```{r}
## species list and number of detections
results_deploy_control |>
  group_by(species) |>
  summarise(num_instances_spp = n()) |>
  pander()

length(unique(results_deploy_control$species)) ## more than half the number of species detected at beaver sites
```


```{r}
## the species and counts of each species recorded at each control deployment (lumped)
spp_numrec_depl_control <- results_deploy_control |>
  group_by(deployment_num, species) |>
  summarise(num_rec = n())
spp_numrec_depl_control

## the species and counts of each species recorded at each control deployment (separated)

## AM0063
results_deploy_control |>
  filter(deployment_num == "AM0067") |>
  group_by(species) |>
  summarise(num_rec = n())

## AM0064
results_deploy_control |>
  filter(deployment_num == "AM0066") |>
  group_by(species) |>
  summarise(num_rec = n())

## AM0081
results_deploy_control |>
  filter(deployment_num == "AM0086") |>
  group_by(species) |>
  summarise(num_rec = n())

## AM0083
results_deploy_control |>
  filter(deployment_num == "AM0089") |>
  group_by(species) |>
  summarise(num_rec = n())

## AM0087
results_deploy_control |>
  filter(deployment_num == "AM0075") |>
  group_by(species) |>
  summarise(num_rec = n())

## AM0094
results_deploy_control |>
  filter(deployment_num == "AM0073") |>
  group_by(species) |>
  summarise(num_rec = n())

## AM0095
results_deploy_control |>
  filter(deployment_num == "AM0074") |>
  group_by(species) |>
  summarise(num_rec = n())

## AM0096
results_deploy_control |>
  filter(deployment_num == "AM0090") |>
  group_by(species) |>
  summarise(num_rec = n())

```

### High probability species detected

```{r}
highprob_results_deploy <- results_deploy_final |>
  filter(probability >= 0.8)

highprob_results_deploy |>
  group_by(species) |>
  summarise(num_instances_spp = n()) |>
  pander()

## the total number of species/sounds recorded across all sites
length(unique(highprob_results_deploy$species))
```



