---
title: "Dirichlet regression"
format: html
editor: visual
---

## Dirichlet regression

Challenge: Liz has data on the proportion of habitat at beaver ponds that falls under different land use categories. Since the data are proportional and should add to 100%, a change (increase or decrease) in one habitat type will automatically indicate a proportional change in other habitat types – thus, the habitat types are dependent.

We can model proporitonal data with a beta regression when we have 2 categories, and Dirichlet regression when the dataset includes more than 2 categories of what is considered *continuous proportional data*. See [this paper](https://besjournals.onlinelibrary.wiley.com/doi/full/10.1111/2041-210X.13234) for details.

## Set up and pull in data

```{r}
rm(list = ls())
library(tidyverse)
library(here)
library(DirichletReg)
library(lmtest)
library(reshape)
library(ggtern)

```

Landcover data

```{r}
lulc <- read.csv(here("Data", "LULC_spillover_diversity.csv"))
glimpse(lulc)
```

To do: \* Convert to long \* filter to just the percentages

```{r}
lulc_narrow <- lulc |>
  select(1, 2, 17:29) 
glimpse(lulc_narrow)
glimpse(lulc)

lulc_combined <- lulc_narrow |>
  mutate(p_Developed = p_Developed_open + p_Developed_low_intensity + p_Developed_med_intensity) |>
  mutate(p_water = p_Open_water + p_Woody_wetlands + p_Emergent_herb_wetlands) |>
  mutate(p_forest = p_Deciduous + p_Evergreen + p_Mixed_forest) |>
  mutate(p_grassy = p_Grassland_herbaceous + p_Pasture_hay + p_Shrub_scrub) |>
  relocate(p_Developed, p_water, p_forest, p_grassy, .after = area_ha) |>
  mutate(treatment = if_else(DEPLOYMENT_NUM == "AM0067" | DEPLOYMENT_NUM == "AM0066" | DEPLOYMENT_NUM == "AM0086" | DEPLOYMENT_NUM == "AM0089" | DEPLOYMENT_NUM == "AM0075" | DEPLOYMENT_NUM == "AM0073" | DEPLOYMENT_NUM == "AM0074" | DEPLOYMENT_NUM == "AM0090" | DEPLOYMENT_NUM == "AM0065",
                             true = "control",
                             false = "beaver")) |>
  relocate(treatment)
  
```

That all came from Liz's code in `high_prop_beaver_biodiv.qmd`. Now make long so that i have a "habitat" variable with the different land use types.

```{r}
lulc_long <- lulc_combined |>
  pivot_longer(cols = p_Developed:p_grassy, names_to = "LU_category", values_to = "percent") |>
  select(treatment, DEPLOYMENT_NUM, area_ha, LU_category, percent)
```

Visualize

```{r}
ggplot(lulc_long, aes(area_ha, percent, color = treatment))+
  geom_point()+
  facet_wrap(~LU_category)+
  coord_flip()
```



The model will break if there are zeros for any of the data, so one solution is to put in a very small number for all the zero values. Let's do that:

```{r}
lulc_long <- lulc_long |>
  mutate(
    new_percent = case_when(
      percent == 0 ~ 0.001,
      percent != 0 ~ percent
    )
  )
```

Now plot again:

```{r}
ggplot(lulc_long, aes(LU_category,  percent, color = treatment))+
  geom_jitter()+
  facet_grid(treatment~area_ha)+
  coord_flip()
```

Now model:
Now fit model. Basically asking does percent of land in each category differ as a function of treatment or scale (1.1 ha vs. 7.7ha  vs. 17.6 ha)

I think we need wide format data for the regression package:

Need a matrix or data frame with nonnegative values of all of the compositional variables.

Use LULC combined

```{r}
lulc_for_reg <- lulc_combined |>
  select(treatment:p_grassy)
```

fill in zeroes

```{r}
lulc_for_reg <- lulc_for_reg |>
  mutate(
    p_Developed = case_when(p_Developed == 0 ~0.001,
                            p_Developed != 0 ~ p_Developed),
    p_water = case_when(p_water == 0 ~0.001, p_water!= 0 ~ p_water),
    p_forest = case_when(p_forest == 0 ~0.001, p_forest!= 0 ~ p_forest),
    p_grassy = case_when(p_grassy == 0 ~0.001, p_grassy != 0 ~ p_grassy)
  )
```
Now prep data:

The 4 proportions serve as the dependent variables and need to be put into a matrix using function `DR_data`:
```{r}
lulc_for_reg$Y <- DR_data(lulc_for_reg[, 4:7], base = 1)
#base = 1 determines the "reference class) -- the component that is not explicitly modeled. 
```
Ok! Now we can fit the model:

```{r}
res1 <-DirichReg(Y ~ treatment + area_ha, data = lulc_for_reg, model = "alternative")
```
Got warnings, so let's try making both treatment and area_ha into factors and see what happens

```{r}
lulc_for_reg$treatment <- as.factor(lulc_for_reg$treatment)
lulc_for_reg$area_ha <- as.factor(lulc_for_reg$area_ha)

res1 <-DirichReg(Y ~ treatment + area_ha, data = lulc_for_reg, model = "alternative")

```
Now view output:

```{r}
summary(res1)
```
Now get plot predicted values for each LULC component along with the observed data
```{r}
#obtain fitted values
pred <- fitted(res1)
pred <- data.frame(lulc_for_reg[, 1:3], pred )
```
Now melt the predicted values (make them long) for use in ggplot
```{r}
pred.melt <- melt(pred, id.var = 1:3)
pred.melt.sim <- unique(pred.melt)

```

Now plot
```{r}
ggplot()+
  geom_point(aes(x = LU_category, percent,color = treatment), data = lulc_long)+
  geom_line(aes(variable, value, group = treatment), data = pred.melt.sim)+
  facet_grid(treatment ~area_ha, scales = "free_y")+
  coord_flip()

ggplot()+
  geom_point(aes(as.numeric(area_ha), percent, color = treatment), data = lulc_long)+
  geom_line(aes(as.numeric(area_ha), value, group = treatment), data = pred.melt.sim)+
  facet_wrap(~LU_category)+
  coord_flip()
```

According to my analysis, there are not significant differences among the different land use classes between treatments or between spatial scales.


