---
title: "elb-jaccard-specnum"
format: html
editor: visual
---

## Set up

```{r}
rm(list=ls())
library(here)
library(tidyverse)
library(vegan)
library(betapart) #turns on package for beta diversity (needed for Jaccard)
library(rstatix) #turns on package with some matrix manipulation functions
```

Read in data

```{r}
jaccard <- read_csv(here("Data/results_wild_mean_per_day.csv"))
```

Get rid of SD and pivot wider

```{r}
jaccard <- jaccard |>
  select(-sd_detections_per_day)|>
  pivot_wider(names_from = species, values_from = mean_detections_per_day)
```

Now get a version where we 1) replace all NA values with 0 and 2)replace all values that are not NA with 1 for presence/absence.

```{r}
#replace na with zero
jaccard_ra <- jaccard
jaccard_ra[is.na(jaccard_ra)] <- 0
  
```

Apparentely, Jaccard should really use a presence/absence matrix, which we will also use for shared species between sites.

```{r}
#requires a dataset of presence/absence data so make that first
PresAb<-jaccard_ra
	
PresAb[, 3:36][PresAb[, 3:36] != 0]<- 1 #assigns all values > 0  to 1 
PresAb
```

Now we have two dataframes, jaccard_ra which includes mean detections per species per day and PresAb, which includes 1 = present and 0 = absent.

What happens if you calculate Jaccard index on a dataframe with relative abundance?

```{r}
Jaccard_rel_abund <- betadiver(jaccard_ra[, 3:38], "j") #j indicates to use the a/a+b+c formula
#result is of class 'distance' - convert to a matrix
Jaccard_rel_abund <- as.matrix(Jaccard_rel_abund)
```

And on a presence-absence dataframe?

```{r}
Jaccard_pres_abs <- betadiver(PresAb[,3:38], "j")
Jaccard_pres_abs <- as.matrix(Jaccard_pres_abs)
```

Are the results the same?

```{r}
Jaccard_pres_abs == Jaccard_rel_abund
```

Yep - gives same result. So package is automatically converting to presence/absense data.

How about species richness? Set up space to hold results

```{r}
div_results <- jaccard_ra[,1:2]
```

Species richness

```{r}
div_results$SR_rel_abundance <- specnumber(jaccard_ra[,3:38]) #using relative abundance in df
div_results$SR_pres_abs <- specnumber(PresAb[,3:38]) #using presence absence
```

Where it should matter is with simpson/shannon

```{r}
div_results$Shannon_rel_abundance <- diversity(jaccard_ra[,3:38], index = "shannon")
div_results$Shannon_pres_absence <- diversity(PresAb[,3:38], index = "shannon")
```

Ok, working as expected.
