---
title: "Ordination Higher prob data - Beaver Biodiv Analysis"
author: Liz Anderson
format: html
---

## Set up
```{r}
rm(list=ls())
library(here)
library(tidyverse)
library(vegan)

#add more libraries
library(ggvegan)
library(ggpubr)
library(ggrepel)

theme_teach<-function(){
  theme(
  #modify text, axis
  axis.text = element_text(color = "black", size = 12),
  axis.title = element_text(color = "black", size = 16)
  )
}
```

```{r}
## set up datasets
results_deploy_final <- read_csv(here("Data", "results_final.csv"))

highprob_results_deploy <- results_deploy_final |>
  filter(probability >= 0.8)

results_wild <- results_deploy_final |>
  filter(species != "Engine" & species != "Dog") ## this will be the dataframe to use

highprob_results_wild <- results_wild |>
  filter(probability >= 0.8)
```


Use the high prob data to compare with ordination plot with lower data from the file `Ordination-pt2-BeaverBiodiv.qmd`



`highprob_results_wild`

## plot detector date ranges

```{r}
## make a column for the numeric day of year
highprob_results_wild$date <- as.Date(highprob_results_wild$date)
highprob_results_wild

highprob_results_wild$day_of_year <- yday(highprob_results_wild$date)
highprob_results_wild
```

## Pivot data wider for use in vegan package

```{r}
library(vegan)

results_summarise_day <- highprob_results_wild |>
  group_by(deployment_num, treatment, species, day_of_year) |>
  summarise(num_obsv = n())
results_summarise_day

results_wide_day <- results_summarise_day |>
  pivot_wider(names_from = species,
              values_from = num_obsv)

results_wide_day[is.na(results_wide_day)] <- 0
results_wide_day


results_summarise <- highprob_results_wild |>
  group_by(deployment_num, treatment, species) |>
  summarise(num_obsv = n())
results_summarise

results_wide <- results_summarise |>
  pivot_wider(names_from = species,
              values_from = num_obsv)

results_wide[is.na(results_wide)] <- 0
results_wide
```

## Ordination work - following along with Ordination-in-class R project

```{r}
SR <- specnumber(results_wide[3:34])

PlotInfo <- results_wide[1:2]

PlotInfo$Shannon <- diversity(results_wide[3:34], index = "shannon")
PlotInfo
```

```{r}
## ordination
BirdOrd<-metaMDS(
  results_wide[, 3:34], #tells vegan which columns just have species info
  distance = "bray", #tells vegan what kind of distance to calculate
  k = 2, #tells vegan that we want 2-d output
  trymax = 50) #tells vegan the number of tries

BirdOrd$stress ## check if the stress value is less than 0.2
stressplot(BirdOrd)

gof<-goodness(BirdOrd) #goodness of fit
plot(BirdOrd, type = "t", main = "goodness of fit")
#points(BirdOrd, display = "sites", cex=gof*100) ## this didn't work we don't need it!

```

Plot results further

__It's not working again__
```{r}
ForPlotting <- fortify(BirdOrd) ## need the extra packages for this to run
```

Ok! That's good!

Now we will make our plots from the new ForPlotting dataframe. \## ggplots \### Make a basic plot

```{r}
ggplot(ForPlotting, aes(nmds1,nmds2,group = score))+
   geom_text(aes(label=label, color = score))+
   scale_color_manual(values = c(sites = "brown", species = "green"))+
   geom_point(aes(color=score, shape =score))+
   scale_color_manual(values = c(sites = "red", species = "black"))
```

This plot is in the direction that we want, but we can improve upon it considerably.

```{r}
ggplot()+
   geom_point(data = subset(ForPlotting, score =="sites"),
              mapping = aes(x = nmds1, y = nmds2),
              alpha = 0.8, 
              size = 4)+
   geom_segment(data=subset(ForPlotting, score == "species"),
                mapping = aes(x = 0, y = 0, xend = nmds1,   yend=nmds2),
                 arrow = arrow(length = unit(0.015, "npc"),
               type = "closed"),
                 color = "darkgray",
                 size = 1,
               alpha = 0.8)+
   geom_text(data = subset(ForPlotting, score == "species"),
             mapping = aes(label = label, x = nmds1*1.1, y=nmds2*1.1), alpha = 0.8)+
   geom_abline(intercept = 0, slope = 0, linetype = "dashed", size = 0.8, color = "gray")+
   geom_vline(aes(xintercept = 0), linetype = "dashed", size = 0.8, color = "gray")+
   theme(
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      panel.background = element_blank(),
      axis.line = element_line(color = "black"),
      legend.position = "none")
   

```

That plot has a lot of information in one plot - let's see if we can pull information out into two separate plots for easier interpretation. Regard the following code which creates a plot called p1 which puts each site on a plot, colored by forest, and numbered by plot number.

### Final plots

```{r}
p1<-ggplot()+
  geom_point(
    data = subset(ForPlotting, score =="sites"),
    mapping = aes(x = nmds1, y = nmds2, color = PlotInfo$treatment),
              alpha = 0.8, size = 4.5)+
  geom_text_repel(
    data = subset(ForPlotting, score =="sites"), 
    mapping = aes(x = nmds1, y = nmds2,label = label))+
  geom_segment(
    data=subset(ForPlotting, score == "species"),
    mapping = aes(x = 0, y = 0, xend = nmds1,   yend=nmds2),
      arrow = arrow(
        length = unit(0.015, "npc"),
        type = "closed"),
      color = "darkgray",
      size = 0,
      alpha = 0)+
  geom_text(
    data = subset(ForPlotting, score == "species"),
    mapping = aes(label = label, x = nmds1*1.1, y=nmds2*1.1), 
      alpha = 0)+
  geom_abline(
    intercept = 0, 
    slope = 0, 
    linetype = "dashed", 
    size = 0.8, color = "gray")+
  geom_vline(
    aes(xintercept = 0), 
    linetype = "dashed", 
    size = 0.8, color = "gray")+
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.line = element_line(color = "black"),
    legend.position = "right",
    legend.title = element_text(size = 20),
    legend.text = element_text(size = 18)) +
  scale_color_manual(values = c("#7fcdbb", "#2c7fb8")) +
  labs(color = "Treatment")
p1
```


And now let's make a second plot that shows the species locations

```{r}
p2<-ggplot()+
  geom_point(
    data = subset(ForPlotting, score =="sites"),
    mapping = aes(x = nmds1, y = nmds2),
    color = "black",
    alpha = 0)+
  geom_segment(
    data=subset(ForPlotting, score == "species"),
    mapping = aes(x = 0, y = 0, xend = nmds1, yend=nmds2),
      arrow = arrow(length = unit(0.015, "npc"),
        type = "closed"),
        color = "darkgray",
        size = 0.8)+
#  geom_text(
#    data = subset(ForPlotting, score == "species"),
 #   mapping = aes(label = label, x = nmds1*1.1, y=nmds2*1.1))+
  geom_text_repel(
    data = subset(ForPlotting, score =="species"), 
    mapping = aes(x = nmds1, y = nmds2,label = label))+
  geom_abline(
    intercept = 0, 
    slope = 0, 
    linetype = "dashed", 
    size = 0.8, 
    color = "gray")+
  geom_vline(aes(xintercept = 0), 
             linetype = "dashed", 
             size = 0.8, 
             color = "gray")+
   theme(
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      panel.background = element_blank(),
      axis.line = element_line(color = "black")
   )
p2
```

Finally, plot them together using `ggpubr`

```{r}
BeaverOrdination<-ggarrange(p1,p2,ncol = 1)
BeaverOrdination

ggsave(here("Output", "HP_BeaverOrdination.jpg"), plot = BeaverOrdination, width = 11.24, height = 10.9, units = "in")


```

















