---
title: "Beaver Biodiversity High Probability"
format: html
---

This document will run through similar code to the the `code_beaver_biodiv.qmd` file but with the filtered data to high BirdNet probability (>0.8)



## Set up

```{r}
rm(list=ls())
library(here)
library(tidyverse)
library(vegan)
```

```{r}
results_deploy_final <- read.csv(here("Data", "results_final.csv"))
```

Scale by 10 detection days and filter
```{r}
results_deploy_final |>
  mutate(date = ymd(date)) |>
  group_by(deployment_num) |>
  summarise(first_rec = min(date),
            last_rec = max(date))

results_deploy_final <- results_deploy_final |>
  mutate(date = ymd(date)) |>
  group_by(deployment_num) |>
  mutate(first_rec = min(date),
         last_rec = max(date))

## mutate a new column with lubridate to get number of deployment days for each - then group and summarize for each detector and species 

results_deploy_final <- results_deploy_final |>
  group_by(deployment_num) |>
  mutate(recording_days = last_rec - first_rec)

glimpse(results_deploy_final)

## to look at these columns:
results_deploy_final |>
  relocate(first_rec, last_rec, recording_days)


## range of recording days:
results_deploy_final |>
  ggplot(aes(x = recording_days))+
  geom_histogram()

results_deploy_final |>
  group_by(treatment) |>
  summarise(
    min_days = min(recording_days),
    max_days = max(recording_days)
  ) |>
  pander()
## looks fairly equal between the treatments
```


__Filter__ to just the rows where the probability is greater than or equal to 0.8
```{r}
highprob_results_deploy <- results_deploy_final |>
  filter(probability >= 0.8)

```

Check the range of recording days again with the high prob data:
```{r}
## range of recording days:
highprob_results_deploy |>
  ggplot(aes(x = recording_days))+
  geom_histogram()

highprob_results_deploy |>
  group_by(treatment) |>
  summarise(
    min_days = min(recording_days),
    max_days = max(recording_days)
  ) |>
  pander()
## same values as not filtered to high prob


## now do the scaling with the filtered data:

highprob_results_deploy <- highprob_results_deploy |>
  mutate(recording_days = as.numeric(recording_days)) |>
  relocate(recording_days)

detections_perdeploy_perspp_highp <- highprob_results_deploy |>
  group_by(deployment_num, species, recording_days, treatment) |>
  summarise(num_detections = n())
  
## get the average number of detections per 10 days
scaled_detections_highp <- detections_perdeploy_perspp_highp |>
  mutate(scaled_num_detections = (num_detections / recording_days) * 10)
scaled_detections_highp


```
### table for poster
```{r}
avg_detections_spp_table_high <- scaled_detections_highp |>
  relocate(treatment) |>
  select(c("treatment", "deployment_num", "species", "scaled_num_detections")) |>
  group_by(treatment, species) |>
  summarise(mean_detections_10days = mean(scaled_num_detections))

avg_detections_spp_table_wide_high <- avg_detections_spp_table_high |>
  pivot_wider(names_from = treatment, 
              values_from = mean_detections_10days)
 
avg_detections_spp_table_wide_high[is.na(avg_detections_spp_table_wide_high)] <- 0

avg_detections_spp_table_wide_rounded_high <- avg_detections_spp_table_wide_high |>
  mutate(across(c(beaver, control), round, digits = 3)) |>
  mutate(Species = species,
         Beaver = beaver,
         Control = control) |>
  select(c("Species", "Beaver", "Control"))

library(gt)

gt_table <- gt(avg_detections_spp_table_wide_rounded_high)

gtsave(gt_table, here("Output", "spp_table.png"))
```




```{r}
## Not scaled
highprob_by_treatment_df <- highprob_results_deploy |>
  group_by(species, treatment) |>
  summarise(num_detections = n()) |>
  pivot_wider(names_from = treatment,
              values_from = num_detections)

## Scaled
highprob_by_treatment_df <- scaled_detections_highp |>
  group_by(species, treatment) |>
  summarise(mean_detections = mean(scaled_num_detections)) |>
  pivot_wider(names_from = treatment,
              values_from = mean_detections)
```

```{r}
highprob_plot_df <- highprob_results_deploy |>
  group_by(deployment_num, species, treatment) |>
  summarise(num_rec = n())
highprob_plot_df

ggplot(data = highprob_plot_df, aes(x = species, y = num_rec, colour = treatment)) +
  geom_segment(aes(x = species, xend = species,
                   y = 0, yend = num_rec)) +
  geom_point() +
  coord_flip() +
  facet_wrap(~deployment_num, scales = "free") +
  theme_bw() +
  labs(x = "Number of recorded detections",
       y = "Species")


## try a diff way to get the scaled
ggplot(data = scaled_detections_highp, aes(x = species, y = scaled_num_detections, colour = treatment, group = treatment)) +
  geom_segment(aes(x = species, xend = species,
                   y = 0, yend = scaled_num_detections)) +
  geom_point() +
  coord_flip() +
  facet_wrap(~deployment_num, scales = "free") +
  theme_bw() +
  labs(y = "Number of detections per 10 days",
       x = "Species")
```
Pivot wider
```{r}
results_wide_high_scaled <- scaled_detections_highp |>
  ungroup() |>
  relocate(recording_days, num_detections) |>
  select(-recording_days, -num_detections) |>
  filter(species != "Engine" & species != "Dog") |>
  pivot_wider(names_from = species,
              values_from = scaled_num_detections)

results_wide_high_scaled[is.na(results_wide_high_scaled)] <- 0

## not needed here:

# results_wide_high_scaled <- results_wide_high_scaled |>
#  mutate(treatment = if_else(deployment_num == "AM0067" | deployment_num == "AM0066" | deployment_num == "AM0086" | deployment_num == "AM0089" | deployment_num == "AM0075" | deployment_num == "AM0073" | deployment_num == "AM0074" | deployment_num == "AM0090" | deployment_num == "AM0065",
#                             true = "control",
#                             false = "beaver")) |>
#  relocate(deployment_num, treatment) 

results_wide_high_scaled
```
remove recording days and num detections

## Diversity indices

```{r}
library(vegan) #turns on package for calculating diversity indices
library(betapart) #turns on package for beta diversity (needed for Jaccard)
library(rstatix) #turns on package with some matrix maniuplation functions


## finish setting up my dataframe to use in vegan (remove the extra columns)
Div_Indices_high <- results_wide_high_scaled[,1:2]

results_scaled_vegan_high <- results_wide_high_scaled[,-(1:2)]
results_scaled_vegan_high

## Species Richness
Div_Indices_high$SR <- specnumber(results_scaled_vegan_high)
Div_Indices_high

## Simpson's Index
Div_Indices_high$Simpson <- diversity(results_scaled_vegan_high, index = "simpson")
Div_Indices_high

## Shannon Index
Div_Indices_high$Shannon <- diversity(results_scaled_vegan_high, index = "shannon")
Div_Indices_high

## summary statistics
# filter to beaver and control so can be compared
Div_Indices_Beaver_high <- Div_Indices_high |>
  filter(treatment == "beaver")

Div_Indices_Control_high <- Div_Indices_high |>
  filter(treatment == "control")

Div_Ind_Summary_all_high <- do.call(cbind, lapply(Div_Indices_high, summary))
Div_Ind_Summary_all_high

Div_Ind_Summary_beaver_high <- do.call(cbind, lapply(Div_Indices_Beaver_high, summary))
Div_Ind_Summary_beaver_high

Div_Ind_Summary_control_high <- do.call(cbind, lapply(Div_Indices_Control_high, summary))
Div_Ind_Summary_control_high

Div_Ind_Summary_beaver_high ## Median values are higher for all indices of the beaver sites
Div_Ind_Summary_control_high
```
Visualize the diversity indices:
```{r}
## pivot to longer format
Div_Long_high <- Div_Indices_high |>
  pivot_longer(cols = SR:Shannon, names_to = "Index", values_to = "Value")


## plot
full_div_plot_high <- ggplot(Div_Long_high, aes(x = fct_reorder(deployment_num, treatment), y = Value, group = Index, color = treatment))+
  geom_point(size = 2) +
  facet_wrap(~Index, scales = "free") +
  labs(x = "Deployment number", color = "Treatment") +
  coord_flip() +
  theme_bw()
full_div_plot_high

# ggsave(here("Output", "HIGH_full_div_plot.jpg"), plot = full_div_plot_high, width = 10, height = 7, units = "in")
```

```{r}
Div_Indices_high |>
  group_by(treatment) |>
  summarise(mean_SR = mean(SR),
            mean_Shannon = mean(Shannon),
            mean_Simpson = mean(Simpson))

SR_plot_high <- ggplot(data = Div_Indices_high, aes(x = treatment, y = SR)) +
  geom_boxplot()+
  geom_point() +
  labs(x = "Treatment", y = "Species Richness") +
  theme_bw()

Simpson_plot_high <- ggplot(data = Div_Indices_high, aes(x = treatment, y = Simpson)) +
  geom_boxplot()+
  geom_point() +
  labs(x = "Treatment", y = "Simpson's Index") +
  theme_bw()

Shannon_plot_high <- ggplot(data = Div_Indices_high, aes(x = treatment, y = Shannon)) +
  geom_boxplot()+
  geom_point() +
  labs(x = "Treatment", y = "Shannon Index") +
  theme_bw()

SR_plot_high
Simpson_plot_high
Shannon_plot_high

# ggsave(here("Output", "HIGH_SR_Plot.jpg"), plot = SR_plot_high, width = 6, height = 11, units = "in")

# ggsave(here("Output", "HIGH_Simpson_Plot.jpg"), plot = Simpson_plot_high, width = 6, height = 11, units = "in")

# ggsave(here("Output", "HIGH_Shannon_Plot.jpg"), plot = Shannon_plot_high, width = 6, height = 11, units = "in")
```

### t-test

```{r}
# "t.test(Ozone ~ Garden.location, data = ozone)
# t.test(response variable ~ predictor variable, data = dataframe)"

t.test(Shannon ~ treatment, data = Div_Indices_high) ## not significant

t.test(Simpson ~ treatment, data = Div_Indices_high) ## not significant

t.test(SR ~ treatment, data = Div_Indices_high) ## not significant
```





## Placeholder for Jaccard Index



## Landcover data

```{r}
lulc <- read.csv(here("Data", "LULC_spillover_diversity.csv"))
glimpse(lulc)
```
To do:
* Convert to long
* filter to just the percentages

```{r}
lulc_long <- lulc |>
  select(1, 2, 17:29) |>
  pivot_longer(cols = 3:15, names_to = "land_type", values_to = "percent_area")
```

Join to Div_Indices dataframe so that can be used in a model with div indices as response
```{r}
Div_Indices_high_joined <- left_join(Div_Indices_high, lulc_long, by = join_by(deployment_num == DEPLOYMENT_NUM))
Div_Indices_high
```

mutate for treatment
pivot 





## Stat tests
_Stat tests to run:_

* test to see if diff in div indices is significantly different -- paired t-test...
* multiple regression model to see what the best predictors of div indices are
   - predictors: percent of land type, beaver vs control, active beavers or not, distance to ___
   



