---
title: "Jaccard Index - Beaver Biodiversity Analysis"
author: Liz Anderson
format: html
---

## Set up
```{r}
rm(list=ls())
library(here)
library(tidyverse)
library(vegan)
library(betapart) #turns on package for beta diversity (needed for Jaccard)
library(rstatix) #turns on package with some matrix manipulation functions
```

```{r}
## set up the data frames

results_deploy_final <- read_csv(here("Data", "results_final.csv"))

highprob_results_deploy <- results_deploy_final |>
  filter(probability >= 0.8)

results_wild <- results_deploy_final |>
  filter(species != "Engine" & species != "Dog")

highprob_results_wild <- results_wild |>
  filter(probability >= 0.8)
```

## Set up file to use in Jaccard code

```{r}
## scale each species by average number of detections per day PER SITE

scale_spp_setup <- results_wild |>
  group_by(deployment_num) |>
  mutate(date = ymd(date)) |>
  mutate(first_rec = min(date),
            last_rec = max(date)) |>
  mutate(recording_days = last_rec - first_rec) |>
  relocate(first_rec, last_rec, recording_days) |>
  mutate(recording_days = as.numeric(recording_days))

scale_spp_long <- scale_spp_setup |>
  group_by(treatment, deployment_num, species, recording_days) |>
  summarise(num_detections = n()) |> # check that there is variation in recording days variable
  group_by(treatment, deployment_num, species) |>
  summarise(av_spp_site = num_detections / recording_days)
## standard deviation still doesn't work

## I think this is what we want for Jaccard index

```

```{r}
## now make wide so it can be used for jaccard/vegan

scale_spp_wide <- scale_spp_long |>
  pivot_wider(names_from = species, values_from = av_spp_site)

scale_spp_wide[is.na(scale_spp_wide)] <- 0

## select function was not working for some reason...

scale_spp_wide_vegan <- scale_spp_wide[,-(1:2)]
```




### Jaccard index

```{r}
Jaccard <- betadiver(scale_spp_wide_vegan, "j") #j indicates to use the a/a+b+c formula
#result is of class 'distance' - convert to a matrix
Jaccard <- as.matrix(Jaccard)
```

_Determine the shared species between sites:_

Now use `betapart`package to get shared species between sites. The function requires a set of presence/absence rather than abundance data so we have to convert our DF dataset to that.

```{r}
#requires a dataset of presence/absence data so make that first
PresAb<-scale_spp_wide_vegan
	
PresAb[][PresAb[] != 0]<- 1 #assigns all values > 0  to 1 
PresAb
```
Now we can run the function to get the number of shared species between each pair of plots.

```{r}
library(betapart)
Beta<-betapart.core(PresAb)#result is a list; we just need one part
Shared<-Beta$shared
## this is little a, the number of species shared between the communities
```

_Create our display matrix_

Now it would be ideal if we could create a matrix with number of shared species ("a") above the diagonal, species richness of each plot along the diagonal, and Jaccard index below the diagonal.

Let's start by adding row and column names to our existing matrices to keep track of things.

```{r}
row.names(Jaccard)<-scale_spp_wide$deployment_num
colnames(Jaccard)<-scale_spp_wide$deployment_num

row.names(Shared)<-scale_spp_wide$deployment_num
colnames(Shared)<-scale_spp_wide$deployment_num

## adds row and column names so can see which communities are being compared
```

Since we want the number of shared species above the diagonal, let's start with the shared matrix. Conveniently, it also has the number of species per plot as its diagonal.

So all we really need to do is put the Jaccard data below the diagonal.

```{r}
#using base r
Shared[lower.tri(Shared)] <- Jaccard[lower.tri(Jaccard, diag = FALSE)]

## puts jaccard below the diagonal

Shared
Shared <- Shared |>
  round(digits = 3)
```

Write matrix and diversity indices to a csv:
```{r}
# write.csv(Shared, here("Output","JaccardMatrix.csv"), row.names = F)

## row.names=F stops an extra column with row names

```











