---
title: "Compiling a Main Data Frame"
format: html
---

Original Code from Aidan Fauth


## Goal:

In the following code, the output files from all the audio moth folders are compiled into a single 'main' data frame.


## Set up

```{r}
rm(list = ls())
library(tidyverse)
library(hms)
library(here)
```


## Compiling and Formatting Data Frame

Here we will take the folder containing all the output files and load them into list:

```{r}
file_list <- list.files(path = here("Data/BirdNet_Output"), pattern = "[0-9].csv$",)

# read in all the files as dfs
df_list <- map(file_list, \(x) read_csv(here("Data/BirdNet_Output", x)))

# process each df to then bind_rows
cleaned_list <- map(df_list, \(x) x |> relocate(file) |>
  pivot_longer(4:ncol(x), names_to = "species", values_to = "probability") |>
  filter(!is.na(probability)))

# add a column with the AM name
for (i in 1:length(cleaned_list)){
  cleaned_list[[i]] <- cleaned_list[[i]] |> 
    mutate(id = str_sub(file_list[[i]], start = 8, end = -5)) # AM names of variable length
}

# compile into single df
temp_df <- bind_rows(cleaned_list)

```

```{r}
table(temp_df$id) ## checking that both beaver and control deployments are included
```


### Create date and time variables

```{r}
# use the "start" variable for the seconds component of the temporal variables 
# (could also use "end", but start makes more sense)
main_df <- temp_df |> mutate(date = str_sub(file, start = 1, end = 8),
          hour = str_sub(file, 10, 11),
          minute = str_sub(file, 12, 13),
          second = start) |> # changed from str_sub(file, 14, 15) bc the file name has no seconds
  unite("date_time", c("date", "hour", "minute", "second"), sep = ":") |>
  mutate(date_time = ymd_hms(date_time),
          time = as_hms(date_time),
         date = as.Date(date_time)) 

## (my addition) added the date only column to the mutate
```


## write csv

```{r}
write_csv(main_df, here("Data/birdnet_results.csv"))
```


## Unite with Audio Moth Deployments

Take a look at the deployment df
```{r}
deploy <- read_csv(here("Data/Beavers-2025.csv"))
glimpse(deploy)
```

### Control Sites

Read in full AM data to then filter to just the control sites. Next bind the control sites with the beaver deployment data to make joining smoother.

```{r}
full_am <- read_csv(here("Data/audio_moth_deployments.csv"))

full_am |>
  glimpse() ## notice two extra columns with no values

## remove the extra columns
full_am <- full_am |>
  select(-"...8", -"...9") |>
  glimpse()


control_deploy <- full_am |>
  filter(deployment_num == "AM0067" | deployment_num == "AM0066" | deployment_num == "AM0086" | deployment_num == "AM0089" | deployment_num == "AM0075" | deployment_num == "AM0073" | deployment_num == "AM0074" | deployment_num == "AM0090" | deployment_num == "AM0065")

glimpse(control_deploy)
## now the two datasets have matching columns so can use bind_rows() to combine

## need to adjust some classes for variables
control_deploy <- control_deploy |>
  mutate(height_above_ground_m = as.numeric(height_above_ground_m))

deploy_combined <- bind_rows(deploy, control_deploy)
```

#### Merge dataframes

```{r}
merged <- left_join(deploy_combined, main_df, join_by("deployment_num" == "id"))
glimpse(merged)
```

```{r}
## make sure it worked by checking how many deployments are present
table(merged$deployment_num)
```




## Save merged data to Data folder

```{r}
write_csv(merged, here("Data/merged_output_deployment.csv"))
```











