---
title: "Multiple regression - beaver biodiv"
author: Liz Anderson
format: html
---

Copied over work from previous qmd (`high_prob_beaver_biodiv.qmd`)





## Set up
```{r}
rm(list=ls())
library(here)
library(tidyverse)
library(vegan)
```

```{r}
## set up the data frames

results_deploy_final <- read_csv(here("Data", "results_final.csv"))

highprob_results_deploy <- results_deploy_final |>
  filter(probability >= 0.8)

results_wild <- results_deploy_final |>
  filter(species != "Engine" & species != "Dog")

highprob_results_wild <- results_wild |>
  filter(probability >= 0.8)
```


To prep data for multiple regression need to:

* scale the data?
* pivot
* Get species richness for each site (the response variable)
* join that with landcover data...



## Landcover data

```{r}
lulc <- read.csv(here("Data", "LULC_spillover_diversity.csv"))
glimpse(lulc)
```


To do:
* Convert to long
* filter to just the percentages

```{r}
lulc_narrow <- lulc |>
  select(1, 2, 17:29) 
glimpse(lulc_narrow)
glimpse(lulc)

lulc_combined <- lulc_narrow |>
  mutate(p_Developed = p_Developed_open + p_Developed_low_intensity + p_Developed_med_intensity) |>
  mutate(p_water = p_Open_water + p_Woody_wetlands + p_Emergent_herb_wetlands) |>
  mutate(p_forest = p_Deciduous + p_Evergreen + p_Mixed_forest) |>
  mutate(p_grassy = p_Grassland_herbaceous + p_Pasture_hay + p_Shrub_scrub) |>
  relocate(p_Developed, p_water, p_forest, p_grassy, .after = area_ha) |>
  mutate(treatment = if_else(DEPLOYMENT_NUM == "AM0067" | DEPLOYMENT_NUM == "AM0066" | DEPLOYMENT_NUM == "AM0086" | DEPLOYMENT_NUM == "AM0089" | DEPLOYMENT_NUM == "AM0075" | DEPLOYMENT_NUM == "AM0073" | DEPLOYMENT_NUM == "AM0074" | DEPLOYMENT_NUM == "AM0090" | DEPLOYMENT_NUM == "AM0065",
                             true = "control",
                             false = "beaver")) |>
  relocate(treatment)
  
```


```{r}
lulc_combined_7.7 <- lulc_combined |>
  filter(area_ha == "7.7")

lulc_combined_17.6 <- lulc_combined |>
  filter(area_ha == "17.6")

Div_Ind_join <- Div_Indices_high |>
  select(c("deployment_num", "SR"))

lulc_joined_7.7 <- left_join(Div_Ind_join, lulc_combined_7.7, by = join_by(deployment_num == DEPLOYMENT_NUM))

lulc_joined_17.6 <- left_join(Div_Ind_join, lulc_combined_17.6, by = join_by(deployment_num == DEPLOYMENT_NUM))


## explore variables to see what predictors to include

ggplot(data = lulc_joined_7.7, aes(x = treatment, y = p_water)) +
  geom_boxplot()
## a bit more in beaver but not much

ggplot(data = lulc_joined_7.7, aes(x = treatment, y = p_Developed)) +
  geom_boxplot()
## more in control

ggplot(data = lulc_joined_7.7, aes(x = treatment, y = p_forest)) +
  geom_boxplot()
## about the same means

ggplot(data = lulc_joined_7.7, aes(x = treatment, y = p_grassy)) +
  geom_boxplot()
## more in beaver

ggplot(data = lulc_joined_7.7, aes(x = treatment, y = p_Woody_wetlands)) +
  geom_boxplot()
## p_woody_wetlands about same
## p_emergent_wetlands more in beaver none in control
```


Add in the active vs not active site data
```{r}
beaver_active_df <- tibble(deployment_num = c("AM0063", "AM0064", "AM0081", "AM0083", "AM0087", "AM0094", "AM0095", "AM0096", "AM0067", "AM0066", "AM0086", "AM0089", "AM0075", "AM0073", "AM0074", "AM0090"), 
       active_yn = c("hist", "hist", "yes", "yes", "yes", "yes", "yes", "hist", "no", "no", "no", "no", "no", "no", "no", "no"))

## paused here because not sure how to distinguish past presence with 
lulc_joined_7.7 <- left_join(lulc_joined_7.7, beaver_active_df, by = join_by(deployment_num))

lulc_joined_7.7 <- lulc_joined_7.7 |>
  relocate(active_yn)

lulc_joined_17.6 <-left_join(lulc_joined_17.6, beaver_active_df, by = join_by(deployment_num)) |>
  relocate(active_yn)
```


### multiple regression
Mult regression model 1 --
beaver predictors 
* treatment
* active
* p_water
* p_grassy

human predictors
* p_developed

```{r}
beaver_model <- lm(data = lulc_joined_7.7, SR ~ treatment + p_water + p_grassy) # activy_yn
summary(beaver_model)

human_model <- lm(data = lulc_joined_7.7, SR ~ p_Developed + treatment)
summary(human_model)


## full model
full_model <- lm(data = lulc_joined_7.7, SR ~ p_Developed + treatment + p_water)
summary(full_model)
```

```{r}
pivot_longer(cols = 3:15, names_to = "land_type", values_to = "percent_area")
```



Join to Div_Indices dataframe so that can be used in a model with div indices as response
```{r}
Div_Indices_high_joined <- left_join(Div_Indices_high, lulc_long, by = join_by(deployment_num == DEPLOYMENT_NUM))
Div_Indices_high
```

mutate for treatment
pivot 






