---
title: "Read-csv-google-drive"
format: html
editor: visual
---

##Overview

When we pull the zooniverse classfication file down, it is huge, and always includes every single classification ever done on our project 🙁

This qmd file helps us to filter to the correct classifcation and workflow version so we get only the data we need.

## Set up workspace and read in CSV file

```{r}
rm(list = ls())
library(tidyverse)
library(here)
library(googledrive)
```

Here's how to read a csv stored in google drive

Create file id to a file that is set to "anyone with the link" on google drive. Here's an example

```{r}
file_id <- "https://drive.google.com/file/d/15KFTBERA-lNsfVv-LMpTC9S3cm_MAZWL/view?usp=sharing"
#need to pass the shared url for the correct csv file here
```

Now get drive info for file

```{r}
my_file <- drive_get(file_id) #pulls info from google drive for the file
my_file$name #prints file name - check to make sure you have pulled correct file
```

Now read file

```{r}
df <-read_csv(drive_read_string(my_file))
```

## Subset to correct workflow and version

The Zooniverse uses different "workflows," each of which can have different versions (as they are tweaked), for classfying images. We start by identifying the workflows present in our classification file.

```{r}
unique(df$workflow_id)
unique(df$workflow_name)
```

The workflow we need to use is called "Species identification"

Let's filter just to that

```{r}
df <- df |> filter(workflow_name == "Species identification")
```

Now we need to see how many versions there are of that worflow

```{r}
unique(df$workflow_version)
```

Generally, we'll want ids from the most current workflow version, in this case 32.13 so we filter to that.

```{r}
df <- df |> filter(workflow_version == 32.13)
```

This reduces us to a much smaller subset of the data.

Next, refer to the qmd called "flatten_species_ids" to see how we get the classification data from the file.

It assumes that you still have `df` in your workspace.