---
title: "Attempting to quantify sampling effort"
author: Liz Anderson
format: html
---



## Set up

```{r}
rm(list=ls())
library(here)
library(tidyverse)
library(vegan)
library(pander)
```

Read in data
```{r}
results_deploy_final <- read_csv(here("Data/results_deploy_final.csv"))
```

## Calculations

* Number of AM days
* Total minutes of recording
* Species detected and how many times total
* Species detected and how many times for each site
* Diversity indices for each site
* Diversity indices for each treatment group

### Number of AM days 

#### All sites
```{r}
## count the number of days
length(unique(results_deploy_final$date))

## number of detections (3 second clips that detected something) on each day
results_deploy_final |>
  count(date)

## total seconds of recording that detected something -- (sum the counts from num observations and multiply by 3)
results_deploy_final |>
  count(date) |>
  summarise(total_times = sum(n)) |>
  mutate(seconds_detecting = total_times * 3) |>
  mutate(minutes_detecting = seconds_detecting / 60) |>
  mutate(hours_detecting = minutes_detecting / 60)

```
## plot detector date ranges

```{r}
## make a column for the numeric day of year
results_deploy_final$date <- as.Date(results_deploy_final$date)
results_deploy_final

results_deploy_final$day_of_year <- yday(results_deploy_final$date)
results_deploy_final
```




```{r}
ggplot(data = results_deploy_final, aes(x = date, color = treatment)) +
  geom_bar() +
  coord_flip() +
  facet_grid(~deployment_num) +
  theme_minimal()
```
First day of a recording 9/21 (sept 21) 
to last day of a recording 12/12 (dec 12)

Recording range for each camera:
```{r}
results_deploy_final |>
  group_by(deployment_num) |>
  summarise(first_rec = min(date),
            last_rec = max(date))

results_deploy_final <- results_deploy_final |>
  group_by(deployment_num) |>
  mutate(first_rec = min(date),
         last_rec = max(date))

## mutate a new column with lubridate to get number of deployment days for each - then group and summarize for each detector and species 

results_deploy_final <- results_deploy_final |>
  group_by(deployment_num) |>
  mutate(recording_days = last_rec - first_rec)
```

What's the range of number of recording days per detector?

```{r}
results_deploy_final |>
  ggplot(aes(x = recording_days))+
  geom_histogram()
```
Quite a range in number of days recording
```{r}
results_deploy_final |>
  group_by(treatment) |>
  summarise(
    min_days = min(recording_days),
    max_days = max(recording_days)
  ) |>
  pander()
```
## Species accumulation
First arrange data by deployment, recording start date, species

```{r}
results_deploy_final <- results_deploy_final |>
  arrange(deployment_num, first_rec, species) 

#View(results_deploy_final)
```

What do I want the final data frame structure to be?

columns for deployment number, date, num_species_today, num_sp_shared_yesterday, num_new_species, cum_num_species

Create placeholder dataframe
```{r}
species_accum <- data.frame(
  deployment_num = character(),
  treatment = character(),
  date = POSIXct(),
  num_sp_today = integer(),
  num_sp_shared_yesterday = integer(),
  num_new_sp = integer(),
  cum_num_species = integer()
)
```

Prepare method to loop through data
```{r}
deployments <- unique(results_deploy_final$deployment_num)
beavers <- c("AM0063","AM0064","AM0081","AM0083", "AM0087","AM0094","AM0095","AM0096")
#empty list to hold lists of species per detector
#create list using deployments so that each list element is named to a particular deployment

sp_list_by_deployment <- sapply(deployments, function(x) NULL)

#second empty list to hold lists of cumulative species per detector
cum_sp_list_by_deployment <- sapply(deployments, function(x) NULL)

#start loop
for(i in 1:length(deployments)){
  my_deployment <- deployments[i]
  if(my_deployment %in% beavers){
    my_treatment <- "beavers"} else {
      my_treatment <- "control"
    }
  temp_df <- results_deploy_final|> filter(deployment_num == my_deployment) |> arrange(deployment_num, date_time)
  
  #second loop to isolate days
  days <- unique(temp_df$date)
  #empty list to hold species lists
  sp_list_by_day <- vector(mode = "list", length = length(days))
  
  #second empty list to hold cumulative species lists
  cum_sp_by_day <- vector(mode = "list", length = length(days))
  
  for(a in 1:length(days)){
    my_date <- days[a]
    
    day_df <- temp_df |> filter(date == my_date)
    num_sp_today <- n_distinct(day_df$species)
    
    ##track which species are listed that day -- for debugging -- delete when working
    sp_for_day <- unique(day_df$species)
    print(sp_for_day)
    ##
    
    sp_list_by_day[[a]] <- unique(day_df$species)
    if (a == 1){
      num_sp_shared_yesterday <- 0} else {
      num_sp_shared_yesterday <- n_distinct(intersect(sp_list_by_day[[a-1]], sp_list_by_day[[a]]))
      }
    if (a == 1){
      num_new_sp <- num_sp_today } else {
        num_new_sp <- length(setdiff(sp_list_by_day[[a]], sp_list_by_day[[a-1]]))
      }
   
     if (a == 1){
      cum_list <- sp_list_by_day[[a]]} else {cum_list <- c(cum_sp_by_day[[a-1]],sp_list_by_day[[a]])}
    #now get rid of duplicates
    cum_list <- unique(cum_list)
    cum_list
    #add species to cumulative list of species by day
    cum_sp_by_day[[a]] <- cum_list
    
    
    cum_num_species <- n_distinct(cum_list)
    
    temp_output <- as.data.frame(list(my_deployment, my_treatment, my_date, num_sp_today, num_sp_shared_yesterday, num_new_sp, cum_num_species))
   
      names(temp_output) <- names(species_accum)

    species_accum <- rbind(species_accum, temp_output)

  }
  
  #now add the species list by day for a single deployment to the list of species lists by deployment
  sp_list_by_deployment[[i]] <- sp_list_by_day
}
```
Now plot species accumulation
 
```{r}
ggplot(species_accum, aes(date, cum_num_species, group = deployment_num, color = treatment))+
  geom_line()+
  labs(x = "Date", y = "Cumulative number of species")+
  theme_bw()+
  theme(legend.position="none")
  
```


