---
title: "Attempting to quantify sampling effort"
author: Liz Anderson
format: html
---



## Set up

```{r}
rm(list=ls())
library(here)
library(tidyverse)
library(vegan)
library(pander)
```

Read in data
```{r}
results_deploy_final <- read_csv(here("Data/results_deploy_final.csv"))
```

## Calculations

* Number of AM days
* Total minutes of recording
* Species detected and how many times total
* Species detected and how many times for each site
* Diversity indices for each site
* Diversity indices for each treatment group

### Number of AM days 

#### All sites
```{r}
## count the number of days
length(unique(results_deploy_final$date))

## number of detections (3 second clips that detected something) on each day
results_deploy_final |>
  count(date)

## total seconds of recording that detected something -- (sum the counts from num observations and multiply by 3)
results_deploy_final |>
  count(date) |>
  summarise(total_times = sum(n)) |>
  mutate(seconds_detecting = total_times * 3) |>
  mutate(minutes_detecting = seconds_detecting / 60) |>
  mutate(hours_detecting = minutes_detecting / 60)

```
## plot detector date ranges

```{r}
## make a column for the numeric day of year
results_deploy_final$date <- as.Date(results_deploy_final$date)
results_deploy_final

results_deploy_final$day_of_year <- yday(results_deploy_final$date)
results_deploy_final
```




```{r}
ggplot(data = results_deploy_final, aes(x = date, color = treatment)) +
  geom_bar() +
  coord_flip() +
  facet_grid(~deployment_num) +
  theme_minimal()
```
First day of a recording 9/21 (sept 21) 
to last day of a recording 12/12 (dec 12)

Recording range for each camera:
```{r}
results_deploy_final |>
  group_by(deployment_num) |>
  summarise(first_rec = min(date),
            last_rec = max(date))

results_deploy_final <- results_deploy_final |>
  group_by(deployment_num) |>
  mutate(first_rec = min(date),
         last_rec = max(date))

## mutate a new column with lubridate to get number of deployment days for each - then group and summarize for each detector and species 

results_deploy_final <- results_deploy_final |>
  group_by(deployment_num) |>
  mutate(recording_days = last_rec - first_rec)
```

What's the range of number of recording days per detector?

```{r}
results_deploy_final |>
  ggplot(aes(x = recording_days))+
  geom_histogram()
```
Quite a range in number of days recording
```{r}
results_deploy_final |>
  group_by(treatment) |>
  summarise(
    min_days = min(recording_days),
    max_days = max(recording_days)
  ) |>
  pander()
```
## Species accumulation
First arrange data by deployment, recording start date, species

```{r}
results_deploy_final <- results_deploy_final |>
  arrange(deployment_num, first_rec, species) 

View(results_deploy_final)
```



