---
title: "Beaver Biodiv - Scaling"
author: Liz Anderson
format: html
---

Beaver Spillover Biodiversity Project

# Document for scaling the detections by day

## Set up
```{r}
rm(list=ls())
library(here)
library(tidyverse)
library(vegan)
```

```{r}
## set up the data frames

results_deploy_final <- read_csv(here("Data", "results_final.csv"))

highprob_results_deploy <- results_deploy_final |>
  filter(probability >= 0.8)

results_wild <- results_deploy_final |>
  filter(species != "Engine" & species != "Dog")

highprob_results_wild <- results_wild |>
  filter(probability >= 0.8)
```

## Scaling!
### Average number of detections per day for each species

Goal - the average number of detections per day for each species at each treatment

Will end up with 2 numbers for each species - means that can also have standard deviations

```{r}
mean_spp_day_beaver <- results_wild |>
  filter(treatment == "beaver") |>
  group_by(date, species) |>
  summarise(n_detections = n()) |>
  complete(date = seq.Date(min(date), max(date), by = "day")) |>
  mutate(n_detections = replace_na(n_detections, 0)) |>
  summarise(mean_spp_day = mean(n_spp),
            sd_spp_day = sd(n_spp),
            treatment = "beaver") |>
  relocate(treatment)
```

```{r}
## this is without anthropogenic sounds

scaling1 <- results_wild |> 
  group_by(treatment, date, species) |>
  summarise(num_detections = n()) |>
  group_by(treatment) |>
  mutate(first_day = min(date),
            last_day = max(date),
            total_day = last_day - first_day) |>
  mutate(total_day = as.numeric(total_day)) |>
  group_by(treatment, species, total_day) |>
  summarise(total_detections = sum(num_detections)) |>
  summarise(average_detections_day = total_detections/total_day,
            sd_detections_day = sd(average_detections_day))

## this is using the total days for each treatment, but should it instead be the total days when the species was detected?
```


```{r}
## try with total days that the species was detected scaling instead -- I don't think this makes as much sense but does give standard deviation values

tentative_av_detections_day <- results_wild |> 
  group_by(treatment, date, species) |>
  summarise(num_detections = n()) |>
  group_by(treatment, species) |>
  summarise(mean_detections = mean(num_detections),
            sd_detections = sd(num_detections)) |>
  arrange(desc(mean_detections))
  

tent_av_detections_day <- ggplot(data = tentative_av_detections_day, aes(x = fct_reorder(species, mean_detections), y = mean_detections)) +
  geom_point(size = 3, aes(color = treatment)) +
  geom_errorbar(aes(ymin = mean_detections - sd_detections, ymax = mean_detections + sd_detections, color = treatment), width = 0.2) +
  theme_bw() +
  theme(legend.position = "none") +
  scale_color_manual(values = c("#7fcdbb", "#2c7fb8")) +
  labs(y = "Mean detections per day",
       x = "Species") +
  coord_flip() +
  facet_wrap(~treatment)

ggsave(here("Output", "mean_detections_day.jpg"), plot = tent_av_detections_day, width = 8, height = 7)
```
#### Comparing the methods
```{r}
tentative_av_detections_day |>
  arrange(treatment, species)

## compared to

scaling1 |>
  arrange(treatment, species)

## takeaway -- tentative_av_detections_day is not great... the numbers in scaling1 seem more accurate, but the issue is it didn't give a standard deviation.
  
```

Plot scaling1
```{r}
scaling1_plot <- ggplot(data = scaling1, aes(x = fct_reorder(species, average_detections_day), y = average_detections_day)) +
  geom_point(size = 3, aes(color = treatment)) +
#  geom_errorbar(aes(ymin = mean_detections - sd_detections, ymax = mean_detections + sd_detections, color = treatment), width = 0.2) +
  theme_bw() +
  theme(legend.position = "none") +
  scale_color_manual(values = c("#7fcdbb", "#2c7fb8")) +
  labs(y = "Mean detections per day",
       x = "Species") +
  coord_flip() +
  facet_wrap(~treatment)

ggsave(here("Output", "scaling1_plot.jpg"), plot = scaling1_plot, width = 8, height = 7)
```


### Average number of species per day per treatment

```{r}
mean_spp_day_beaver <- results_wild |>
  filter(treatment == "beaver") |>
  group_by(date, species) |>
  summarise() |>
  group_by(date) |>
  summarise(n_spp = n()) |>
  complete(date = seq.Date(min(date), max(date), by = "day")) |>
  mutate(n_spp = replace_na(n_spp, 0)) |>
  summarise(mean_spp_day = mean(n_spp),
            sd_spp_day = sd(n_spp),
            treatment = "beaver") |>
  relocate(treatment)

## Control
mean_spp_day_control <- results_wild |>
  filter(treatment == "control") |>
  group_by(date, species) |>
  summarise() |>
  group_by(date) |>
  summarise(n_spp = n()) |>
  complete(date = seq.Date(min(date), max(date), by = "day")) |>
  mutate(n_spp = replace_na(n_spp, 0)) |>
  summarise(mean_spp_day = mean(n_spp),
            sd_spp_day = sd(n_spp),
            treatment = "control") |>
  relocate(treatment)

## join the two dataframes/treatments

mean_spp_day <- bind_rows(mean_spp_day_beaver, mean_spp_day_control)

```


```{r}
plot_av_spp_day <- ggplot(data = mean_spp_day, aes(x = treatment, y = mean_spp_day)) +
  geom_point(size = 3, aes(color = treatment)) +
  geom_errorbar(aes(ymin = mean_spp_day - sd_spp_day, ymax = mean_spp_day + sd_spp_day, color = treatment), width = 0.2) +
  theme_bw() +
  theme(legend.position = "none") +
  scale_color_manual(values = c("#7fcdbb", "#2c7fb8")) +
  labs(y = "Mean species detected per day",
       x = "Treatment")

ggsave(here("Output", "mean_numspp_perday.jpg"), plot = plot_av_spp_day, width = 7, height = 5)
```








