---
title: "Beaver Biodiv - Scaling"
author: Liz Anderson
format: html
---

Beaver Spillover Biodiversity Project

# Document for scaling the detections by day

## Set up
```{r}
rm(list=ls())
library(here)
library(tidyverse)
library(vegan)
```

```{r}
## set up the data frames

results_deploy_final <- read_csv(here("Data", "results_final.csv"))

highprob_results_deploy <- results_deploy_final |>
  filter(probability >= 0.8)

results_wild <- results_deploy_final |>
  filter(species != "Engine" & species != "Dog")

highprob_results_wild <- results_wild |>
  filter(probability >= 0.8)
```

## Scaling!
### Average number of detections per day for each species

Goal - the average number of detections per day for each species at each treatment

Using the dataframe that is just the wild species (removed engine and dog)

Will end up with 2 numbers for each species - means that can also have standard deviations

```{r}
#### this code is a work in progress

mean_spp_day_beaver <- results_wild |>
  filter(treatment == "beaver") |>
  group_by(date, species) |>
  summarise(n_detections = n()) |>
#  complete(date = seq.Date(min(date), max(date), by = "day")) |> ## this line doesn't work because there is more than 1 row for each date
#  mutate(n_detections = replace_na(n_detections, 0)) |>
  group_by(species) |>
  summarise(mean_detections_spp_day = mean(n_detections),
            sd_detections_spp_day = sd(n_detections),
            treatment = "beaver") |>
  relocate(treatment)

## I think this is the mean based on the number of days when that species was actually detected. It doesn't include the days when that species was not detected, which is important to have for the average. 
```

```{r}
scaling1 <- results_wild |> 
  group_by(treatment, date, species) |>
  summarise(num_detections = n()) |>
  group_by(treatment) |>
  mutate(first_day = min(date),
            last_day = max(date),
            total_day = last_day - first_day) |>
  mutate(total_day = as.numeric(total_day)) |>
  group_by(treatment, species, total_day) |>
  summarise(total_detections = sum(num_detections)) |>
  summarise(average_detections_day = total_detections/total_day,
            sd_detections_day = sd(average_detections_day))

## this is using the total days for each treatment, which I think makes sense. However, it doesn't give a standard deviation. 

```


```{r}
## this is similar to above where it wasn't quite working

## try with total days that the species was detected scaling instead -- I don't think this makes as much sense but does give standard deviation values

tentative_av_detections_day <- results_wild |> 
  group_by(treatment, date, species) |>
  summarise(num_detections = n()) |>
  group_by(treatment, species) |>
  summarise(mean_detections = mean(num_detections),
            sd_detections = sd(num_detections)) |>
  arrange(desc(mean_detections))
  

tent_av_detections_day <- ggplot(data = tentative_av_detections_day, aes(x = fct_reorder(species, mean_detections), y = mean_detections)) +
  geom_point(size = 3, aes(color = treatment)) +
  geom_errorbar(aes(ymin = mean_detections - sd_detections, ymax = mean_detections + sd_detections, color = treatment), width = 0.2) +
  theme_bw() +
  theme(legend.position = "none") +
  scale_color_manual(values = c("#7fcdbb", "#2c7fb8")) +
  labs(y = "Mean detections per day",
       x = "Species") +
  coord_flip() +
  facet_wrap(~treatment)

# ggsave(here("Output", "mean_detections_day.jpg"), plot = tent_av_detections_day, width = 8, height = 7)
```
#### Comparing the methods
```{r}
tentative_av_detections_day |>
  arrange(treatment, species)

## compared to

scaling1 |>
  arrange(treatment, species)

## takeaway -- tentative_av_detections_day is not great... the numbers in scaling1 seem more accurate, but the issue is it didn't give a standard deviation.
  
```

Plot scaling1
```{r}
scaling1_plot <- ggplot(data = scaling1, aes(x = fct_reorder(species, average_detections_day), y = average_detections_day)) +
  geom_point(size = 3, aes(color = treatment)) +
#  geom_errorbar(aes(ymin = mean_detections - sd_detections, ymax = mean_detections + sd_detections, color = treatment), width = 0.2) +
  theme_bw() +
  theme(legend.position = "none") +
  scale_color_manual(values = c("#7fcdbb", "#2c7fb8")) +
  labs(y = "Mean detections per day",
       x = "Species") +
  coord_flip() +
  facet_wrap(~treatment)

# ggsave(here("Output", "scaling1_plot.jpg"), plot = scaling1_plot, width = 8, height = 7)
```


### Average number of species per day per treatment

```{r}
## using the code with function complete() worked well here. I filtered to each treatment to just have one line for each day because I'm counting the number of species each day. Then join the two treatment data frames.

mean_spp_day_beaver <- results_wild |>
  filter(treatment == "beaver") |>
  group_by(date, species) |>
  summarise() |>
  group_by(date) |>
  summarise(n_spp = n()) |>
  complete(date = seq.Date(min(date), max(date), by = "day")) |>
  mutate(n_spp = replace_na(n_spp, 0)) |>
  summarise(mean_spp_day = mean(n_spp),
            sd_spp_day = sd(n_spp),
            treatment = "beaver") |>
  relocate(treatment)

## Control
mean_spp_day_control <- results_wild |>
  filter(treatment == "control") |>
  group_by(date, species) |>
  summarise() |>
  group_by(date) |>
  summarise(n_spp = n()) |>
  complete(date = seq.Date(min(date), max(date), by = "day")) |>
  mutate(n_spp = replace_na(n_spp, 0)) |>
  summarise(mean_spp_day = mean(n_spp),
            sd_spp_day = sd(n_spp),
            treatment = "control") |>
  relocate(treatment)

## join the two dataframes/treatments

mean_spp_day <- bind_rows(mean_spp_day_beaver, mean_spp_day_control)

```


```{r}
plot_av_spp_day <- ggplot(data = mean_spp_day, aes(x = treatment, y = mean_spp_day)) +
  geom_point(size = 3, aes(color = treatment)) +
  geom_errorbar(aes(ymin = mean_spp_day - sd_spp_day, ymax = mean_spp_day + sd_spp_day, color = treatment), width = 0.2) +
  theme_bw() +
  theme(legend.position = "none") +
  scale_color_manual(values = c("#7fcdbb", "#2c7fb8")) +
  labs(y = "Mean species detected per day",
       x = "Treatment")

# ggsave(here("Output", "mean_numspp_perday.jpg"), plot = plot_av_spp_day, width = 7, height = 5)
```

#### stat test compare
To be able to compare these with a Wilcoxon test I think it needs to have multiple points, so I should get the average species per day at each deployment site.

```{r}

site_mean_spp_day_beaver <- results_wild |>
  filter(treatment == "beaver") |>
  group_by(deployment_num, date, species) |>
  summarise() |>
  group_by(deployment_num, date) |>
  summarise(n_spp = n()) |>
  complete(date = seq.Date(min(date), max(date), by = "day")) |> ## I'm surprised this worked because I thought with multiple rows per date it wouldn't
  mutate(n_spp = replace_na(n_spp, 0)) |>
  summarise(mean_spp_day = mean(n_spp),
            sd_spp_day = sd(n_spp),
            treatment = "beaver") |>
  relocate(treatment)

## Control
site_mean_spp_day_control <- results_wild |>
  filter(treatment == "control") |>
  group_by(deployment_num, date, species) |>
  summarise() |>
  group_by(deployment_num, date) |>
  summarise(n_spp = n()) |>
  complete(date = seq.Date(min(date), max(date), by = "day")) |>
  mutate(n_spp = replace_na(n_spp, 0)) |>
  summarise(mean_spp_day = mean(n_spp),
            sd_spp_day = sd(n_spp),
            treatment = "control") |>
  relocate(treatment)

## join the two dataframes/treatments

mean_spp_day_site <- bind_rows(site_mean_spp_day_beaver, site_mean_spp_day_control)
mean_spp_day_site

mean_spp_day_site |>
  group_by(treatment) |>
  summarise(mean_all = mean(mean_spp_day)) ## these values are not the same as the mean I got earlier when didn't group by treatment first... that probably makes a difference for the t-test?
```
Now run a Wilcoxon test to see if distributions for beaver and control site mean number of species detected per day is different

```{r}
wilcox.test(mean_spp_day ~ treatment, data = mean_spp_day_site) ## not significant

t.test(mean_spp_day ~ treatment, data = mean_spp_day_site) ## because this is continuous data, is a t.test appropriate? it's also not significant
```








