---
title: "Beaver Biodiv -- comparing treatments"
author: Liz Anderson
format: html
---

Document for running statistical tests to compare biodiversity between beaver and control sites.

Previously did a t-test like this: `t.test(SR ~ treatment, data = Div_Indices_high)`

But t-tests aren't supposed to be done with count data (which SR is)

## Set up
```{r}
rm(list=ls())
library(here)
library(tidyverse)
library(vegan)
```

```{r}
## set up datasets
results_deploy_final <- read_csv(here("Data", "results_final.csv"))

highprob_results_deploy <- results_deploy_final |>
  filter(probability >= 0.8)

results_wild <- results_deploy_final |>
  filter(species != "Engine" & species != "Dog")

highprob_results_wild <- results_wild |>
  filter(probability >= 0.8)
```

Use this code: `wilcox.test(values ~ groups, data = df)` to run a non-paired Wilcoxon test - does not assume that the data is normally distributed



Using the dataset without anthropogenic noise


### Species richness per site and treatment


```{r}
results_wild_wide <- results_wild |>
  group_by(treatment, deployment_num, species) |>
  summarise(num_detections = n()) |>
  pivot_wider(names_from = species, values_from = num_detections)


results_wild_wide[is.na(results_wild_wide)] <- 0

results_wild_vegan <- results_wild_wide |>
  ungroup() |>
  select(-treatment, -deployment_num)
```

```{r}
Div_Indices_notscaled <- results_wild_wide[,1:2]
Div_Indices_notscaled$SR <- specnumber(results_wild_vegan) 
Div_Indices_notscaled ## this is giving the same +2 values for SR...
```
#### Wilcoxon Test

Now that I have the species richness I can run the Wilcoxon test
```{r}
wilcox.test(SR ~ treatment, data = Div_Indices_notscaled) ## just significant enough -- p-value 0.06429

t.test(SR ~ treatment, data = Div_Indices_notscaled)
```
#### Plot
Option 1 - boxplot, doesn't show means but shows distibution
Could add a point for mean
```{r}
overall_mean_SR <- Div_Indices_notscaled |>
  group_by(treatment) |>
  summarise(mean_overall_SR = mean(SR),
            sd_overall_SR = sd(SR))
```


Option 2 - could do what I did for another plot and have each site as a point, the means as a point and error bars for the means
```{r}
## Option 1
SR_plot_and_mean <- 
  ggplot(data = Div_Indices_notscaled, aes(x = treatment, y = SR, fill = treatment)) +
  geom_boxplot()+
  geom_point() +
  geom_point(data = overall_mean_SR, aes(x = treatment, y = mean_overall_SR), shape = 21, fill = "#edf8b1", color = "black", size = 5) +
  labs(x = "Treatment", y = "Species Richness") +
  theme_bw() +
  coord_flip() +
  scale_fill_manual(values = c("#7fcdbb", "#2c7fb8")) +
  theme(legend.position = "none") +
  scale_y_continuous(breaks = seq(0, 12, by = 1))
  
ggsave(here("Output", "SR_plot_and_mean.jpg"), plot = SR_plot_and_mean, width = 11.25, height = 4.5, units = "in")
```

```{r}
## option 2

ggplot(data = overall_mean_SR, aes(x = treatment, y = mean_overall_SR)) +
  geom_errorbar(data = overall_mean_SR, aes(ymin = mean_overall_SR - sd_overall_SR, ymax = mean_overall_SR + sd_overall_SR, color = treatment), width = 0.2, linewidth = 1) +
  geom_point(shape = 21, fill = "#edf8b1", colour = "black", size = 6, stroke = 1) +
  geom_point(data = Div_Indices_notscaled, aes(x = treatment, y = SR), size = 2) +
  scale_color_manual(values = c("#7fcdbb", "#2c7fb8")) +
  theme_bw()+
  theme(legend.position = "none") +
  labs(x = "Treatment", y = "Species richness per site")
```

#### Mann-Whitney U test

```{r}
wilcox.test(SR ~ treatment, data = Div_Indices_notscaled)


```



### Spp richness High conf
```{r}
highprob_results_wild_wide <- highprob_results_wild |>
  group_by(treatment, deployment_num, species) |>
  summarise(num_detections = n()) |>
  pivot_wider(names_from = species, values_from = num_detections)


highprob_results_wild_wide[is.na(highprob_results_wild_wide)] <- 0

highprob_results_wild_wide

HP_results_wild_vegan <- highprob_results_wild_wide |>
  ungroup() |>
  select(-treatment, -deployment_num)
```

```{r}
HP_Div_Indices_notscaled <- highprob_results_wild_wide[,1:2]
HP_Div_Indices_notscaled$SR <- specnumber(HP_results_wild_vegan) 
HP_Div_Indices_notscaled ## this is giving the same +2 values for SR compared to the Jaccard
```
```{r}
HP_Div_Indices_notscaled |>
  group_by(treatment) |>
  summarise(mean_overall = mean(SR),
            sd_overall = sd(SR))
```


#### Test
```{r}
wilcox.test(SR ~ treatment, data = HP_Div_Indices_notscaled) ## 0.1122

t.test(SR ~ treatment, data = HP_Div_Indices_notscaled) ## 0.067
```





## Species Richness tables 

Get a table to put in paper with the species richness of each site with a column for beaver and a column for control (and maybe also separated by confidence score? - although doesn't change much)

```{r}
Div_Indices

Div_Indices |>
  group_by(treatment) |>
  summarise(mean_SR = mean(SR),
            sd_SR = sd(SR))

write_csv(Div_Indices, here("Output", "SR_all.csv"))




Highprob_Div_Indices |>
  group_by(treatment) |>
  summarise(mean_SR = mean(SR),
            sd_SR = sd(SR))
```

## Species Richness scaled per day

### is there a difference in the number of species detected per day between treatments?

```{r}
scaled_numspp_day <- read_csv(here("Output", "num_spp_by_date_means.csv"))
```
```{r}
scaled_numspp_day
```
```{r}
wilcox.test(mean_num_species_per_day ~ treatment, data = scaled_numspp_day)
## p = 0.2698

t.test(mean_num_species_per_day ~ treatment, data = scaled_numspp_day)

## both not significant at 0.5
```

Plot to visualize a trend

```{r}
grand_mean_num_day_treament <- num_sp_by_date_means |>
  group_by(treatment) |>
  summarise(mean_sp_treatment = mean(mean_num_species_per_day),
            sd_sp_treatment = sd(mean_num_species_per_day))

plot_meanspp_day_treatment <- ggplot(data = grand_mean_num_day_treament, aes(x = treatment, y = mean_sp_treatment)) +
  geom_errorbar(data = grand_mean_num_day_treament, aes(ymin = mean_sp_treatment - sd_sp_treatment, ymax = mean_sp_treatment + sd_sp_treatment, color = treatment), width = 0.2, linewidth = 1) +
  geom_point(shape = 21, fill = "#edf8b1", colour = "black", size = 6, stroke = 1) +
  geom_point(data = scaled_numspp_day, aes(x = treatment, mean_num_species_per_day), size = 2) +
  scale_color_manual(values = c("#7fcdbb", "#2c7fb8")) +
  theme_bw()+
  theme(legend.position = "none") +
  labs(x = "Treatment", y = "Mean number of species detected per day")

ggsave(here("Output", "mean_spp_day_treatment.png"), plot = plot_meanspp_day_treatment, width = 7, height = 5.5)
```





