---
title: "Beaver Biodiv -- comparing treatments"
author: Liz Anderson
format: html
---

Document for running statistical tests to compare biodiversity between beaver and control sites.

Previously did a t-test like this: `t.test(SR ~ treatment, data = Div_Indices_high)`

But t-tests aren't supposed to be done with count data (which SR is)

## Set up
```{r}
rm(list=ls())
library(here)
library(tidyverse)
library(vegan)
```

```{r}
## set up datasets
results_deploy_final <- read_csv(here("Data", "results_final.csv"))

highprob_results_deploy <- results_deploy_final |>
  filter(probability >= 0.8)

results_wild <- results_deploy_final |>
  filter(species != "Engine" & species != "Dog")

highprob_results_wild <- results_wild |>
  filter(probability >= 0.8)
```

Use this code: `wilcox.test(values ~ groups, data = df)` to run a non-paired Wilcoxon test - does not assume that the data is normally distributed

First need the data in a wide format for vegan - and need it scaled with the deployment sites

Using the dataset without anthropogenic noise

```{r}
scale_for_wide <- results_wild |> 
  group_by(treatment, deployment_num, date, species) |>
  summarise(num_detections = n()) |>
  group_by(treatment, deployment_num) |>
  mutate(first_day = min(date),
            last_day = max(date),
            total_day = last_day - first_day) |>
  mutate(total_day = as.numeric(total_day)) |>
  group_by(treatment, deployment_num, species, total_day) |>
  summarise(total_detections = sum(num_detections)) |>
  summarise(average_detections_day = total_detections/total_day)

scaled_day_wide <- scale_for_wide |>
  pivot_wider(names_from = species, values_from = average_detections_day)

scaled_day_wide[is.na(scaled_day_wide)] <- 0

scaled_day_wide
```

```{r}
## set up dataframe
Div_Indices <- scaled_day_wide[,1:2]
Div_Indices$SR <- specnumber(scaled_day_wide)
Div_Indices
```

Now that I have the species richness I can run the Wilcoxon test
```{r}
wilcox.test(SR ~ treatment, data = Div_Indices)
```


## Run it again for 0.8 confidence score

highprob_results_wild

```{r}
highprob_scale_for_wide <- highprob_results_wild |> 
  group_by(treatment, deployment_num, date, species) |>
  summarise(num_detections = n()) |>
  group_by(treatment, deployment_num) |>
  mutate(first_day = min(date),
            last_day = max(date),
            total_day = last_day - first_day) |>
  mutate(total_day = as.numeric(total_day)) |>
  group_by(treatment, deployment_num, species, total_day) |>
  summarise(total_detections = sum(num_detections)) |>
  summarise(average_detections_day = total_detections/total_day)

highprob_scaled_day_wide <- highprob_scale_for_wide |>
  pivot_wider(names_from = species, values_from = average_detections_day)

highprob_scaled_day_wide[is.na(highprob_scaled_day_wide)] <- 0

highprob_scaled_day_wide
```

```{r}
## set up dataframe
Highprob_Div_Indices <- highprob_scaled_day_wide[,1:2]
Highprob_Div_Indices$SR <- specnumber(highprob_scaled_day_wide)
Highprob_Div_Indices
```

```{r}
wilcox.test(SR ~ treatment, data = Highprob_Div_Indices)
```




